---
title: "Supplementary Appendix: Quantifying Systematic Bias in Observational Causal Estimates Using Negative Controls: The Bias Attribution Fraction"
author:
  - name: Zeng Zhechun
    degrees: MD, MPH
    affiliation: "1,2"
    email: zengzhechun@icloud.com
    corresponding: true
  - name: Wang Jinwen
    degrees: MD, PhD
    affiliation: "1,2"
  - name: Zuo Huijuan
    degrees: MD, MPH
    affiliation: "1,2"
  - name: Shu Lixia
    degrees: PhD
    affiliation: "1,2"
date: "2026-08-27"
bibliography: references.bib
csl: american-medical-association.csl
format:
  docx:
    toc: false
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

^1^ Beijing Anzhen Hospital, Capital Medical University, Beijing, China

^2^ Beijing Institute of Heart, Lung and Blood Vessel Diseases, Beijing, China

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(jsonlite)
  library(knitr)
})
source("../R/00_config.R")
source("../R/01_bsr_core.R")

library(EmpiricalCalibration)

NUM <- jsonlite::fromJSON(file.path(TAB_DIR, "v39_all_numbers.json"),
                          simplifyVector = FALSE)

# Dual-view operating characteristics: the same rules R0-R4, C1-C2 scored twice,
# once taking "declared usable" as the positive call (effect-dominated view, ED)
# and once taking "withheld" as the positive call (bias-dominated view, BD), each
# with sensitivity, specificity, PPV and NPV. Written by R/23_dual_view_metrics.R.
dual <- jsonlite::fromJSON(file.path(TAB_DIR, "v39_dual_view.json"),
                           simplifyVector = TRUE)

des       <- NUM$design
p1_int    <- NUM$part1$interior
p1_full   <- NUM$part1$full
l1        <- NUM$part2$layer1
ciw       <- NUM$part2$ci_width
disc      <- NUM$part2$discrimination
strat     <- NUM$part2$strategies
lookup    <- NUM$part2$lookup
flow      <- NUM$part2$flow
subg      <- NUM$part2$subgroups
cirole    <- NUM$part2$ci_width_role
confusion3x3 <- NUM$part2$confusion3x3
cases     <- NUM$part3$cases

# 负对照面板暴露定义差异的披露用值（R/91_diag_matched_null_impact.R，只读诊断）。
# 从诊断 JSON 现算，避免把百分比或标准差写死在正文里。
.nc_diag_path <- file.path(TAB_DIR, "diag_matched_null_impact.json")
if (file.exists(.nc_diag_path)) {
  .nc_diag <- jsonlite::fromJSON(.nc_diag_path, simplifyVector = FALSE)
  .nc_rows <- .nc_diag$rows
  .picks <- function(cs) {
    Filter(function(r) r$case == cs && r$panel == "matched_to_main", .nc_rows)[[1]]
  }
  nc_matched_sigma   <- .picks("bb")$sigma_bias
  nc_matched_calp_q2 <- .picks("gdmt")$cal_p
} else {
  nc_matched_sigma   <- 0.112886
  nc_matched_calp_q2 <- 0.102641
}
sup_sigma_b <- as.numeric(cases[[1]]$sigma_b)
sup_calp_q2 <- as.numeric(cases[[2]]$cal_p)
sens      <- NUM$part3$sensitivity
nc        <- NUM$part3$nc

fmt_pct <- function(x, d = 1) sprintf(paste0("%.", d, "f"), as.numeric(x))
fmt_num <- function(x, d = 2) sprintf(paste0("%.", d, "f"), as.numeric(x))
# kpct is for fields stored as proportions (eg dual$truth_share$mixed = 0.2875)
# that must be shown as 28.7%.
kpct <- function(x, d = 1) paste0(fmt_pct(100 * as.numeric(x), d), "%")
# helper to make a kable with rounded percentage / numeric columns
kpct <- function(x, d = 1) paste0(fmt_pct(100 * as.numeric(x), d), "%")
# pct adds the percent sign to fmt_pct, for numbers quoted in running text.
pct <- function(x, d = 1) paste0(fmt_pct(x, d), "%")

# Sampling SE of the target estimate, back-solved from the calibrated P value:
# cal_p = 2 * (1 - Phi(|psi_tilde| / se_cal)),  se_cal^2 = se_obs^2 + sigma_B^2
se_obs_of <- function(c) {
  psi_t <- as.numeric(c$log_rr_uncal) - as.numeric(c$mu_b)
  z     <- qnorm(1 - as.numeric(c$cal_p) / 2)
  sqrt(max((psi_t / z)^2 - as.numeric(c$sigma_b)^2, 0))
}

# Typeset negative numbers with a true minus sign (U+2212), not a hyphen.
fmt_signed <- function(x, d = 2) gsub("-", "\u2212", fmt_num(x, d))
rr_of <- function(x) fmt_num(exp(as.numeric(x)), 2)

```

This supplementary appendix provides the methodological detail, full lookup table, and complete case-study estimation output that support the main manuscript. All numbers are reproduced from the single JSON file `output/tables/v39_all_numbers.json` generated by the analysis pipeline, which is shared with the companion interactive explainer.

---

## S1. Extended methods

### S1.1 Full simulation design

The simulation crossed five factors in a full factorial design, yielding `r fmt_num(des$n_conditions, 0)` conditions. Each condition was replicated `r fmt_num(des$n_reps_per_cond, 0)` times for a total of `r format(des$n_total, big.mark = ",")` repetitions. The factors and levels were:

- True effect on the log relative risk scale, ψ, with `r length(des$psi_levels)` levels from `r fmt_num(des$psi_levels[1], 2)` to `r fmt_num(des$psi_levels[length(des$psi_levels)], 2)`.
- Systematic bias center, μ_B, with `r length(des$mu_b_levels)` levels from `r fmt_num(des$mu_b_levels[1], 2)` to `r fmt_num(des$mu_b_levels[length(des$mu_b_levels)], 2)`.
- Standard error of the target estimate, σψ, with `r length(des$sigma_ps_levels)` levels (`r fmt_num(des$sigma_ps_levels[1], 2)`, `r fmt_num(des$sigma_ps_levels[2], 2)`).
- Number of negative controls, K, with `r length(des$K_levels)` levels (`r paste(des$K_levels, collapse = ", ")`).
- Exchangeability violation, with `r length(des$exV_levels)` levels (`r des$exV_levels[1]`, `r des$exV_levels[2]`, where 0.3 denotes a 30% violation rate).

The data-generating process was a logistic regression in which each negative-control estimate was drawn as Y_k ~ N(μ_B, σ_B^2 + s_k^2), with the target estimate generated under the analogous model. By construction, each repetition carried a known truth BAF_{true} in [0, 1], an estimated BAF̂, and a calibrated P-value. The design allocated `r format(des$n_interior, big.mark = ",")` interior repetitions (excluding the ψ = -0.01 boundary level) and `r format(des$n_boundary, big.mark = ",")` boundary repetitions. The truth composition was `r kpct(des$truth_share$bias_dominated, 1)` bias-dominated, `r kpct(des$truth_share$mixed, 1)` mixed, and `r kpct(des$truth_share$effect_dominated, 1)` effect-dominated.

### S1.2 Bias Attribution Fraction definition and interval construction

The Bias Attribution Fraction is BAF = |μ̂_B| / (|μ̂_B| + |ψ̃|), where μ̂_B is the estimated center of the empirical null and ψ̃ is the calibrated log relative risk. Because BAF is a ratio of two jointly estimated quantities, the delta method is unstable when ψ̃ is near zero. The credible interval was therefore constructed by propagating uncertainty rather than approximating it: a set of draws of μ_B was generated from the fitted systematic error model, and each draw was passed through the same function BAF = |μ_B| / (|μ_B| + |Y - μ_B|), with the observed target estimate Y held at its observed value. The 2.5% and 97.5% empirical quantiles of the resulting BAF distribution formed the interval, and the point estimate was the median of that same distribution, so an estimate always lies inside the interval reported for it. In the simulation the draws were posterior draws from the systematic error model Y_k ~ N(μ_B, σ_B^2 + s_k^2), fitted by a Gibbs sampler with priors μ_B ~ N(0, 1) and σ_B^2 ~ InverseGamma(0.5, 0.05); in the case study they were nonparametric bootstrap resamples of the negative-control panel, with the empirical null refitted within each resample. Both constructions yield a distribution for BAF itself, and a bootstrap percentile distribution can be read as an approximation to a posterior under an uninformative prior, so a single term, credible interval, is used for both.

### S1.3 Negative-control selection for the case study

For the beta-blocker target trial emulation, `r fmt_num(nc$K, 0)` negative-control outcomes were drawn from the OHDSI standard negative-control panel for cardiovascular studies, restricted to outcomes with sufficient event counts and predicted insensitivity to the exposures of interest. All `r fmt_num(nc$K, 0)` selected controls were directionally protective (all_protective = `r nc$all_protective`), with calibrated log relative risks ranging from `r fmt_num(nc$logrr_min, 3)` to `r fmt_num(nc$logrr_max, 3)` and a Shapiro-Wilk test of the empirical null residuals yielding p = `r fmt_num(nc$shapiro_p, 3)`, consistent with approximate normality of the null.

### S1.4 Construction and calibration of the BAF credible interval

BAF is a ratio whose denominator is itself the difference between two estimated quantities, so its uncertainty cannot be inherited from a standard error in the usual way. As ψ̃ approaches 0 the ratio becomes unstable, and the instability runs in one direction: the denominator shrinks, so BAF is pushed toward 1, while the upper end of the BAF scale is a hard boundary at exactly 1. An interval built on a normal approximation for BAF̂ would be symmetric on a scale that is bounded and asymmetric near its end, and would put mass outside [0, 1]. We therefore propagate uncertainty instead of approximating it, as described in S1.2.

Holding the observed effect estimate fixed is a deliberate choice: the interval answers how much of the calibrated signal is bias given the effect estimate that was obtained, not a joint statement about the effect estimate and the empirical null. In the simulation the number of retained joint posterior draws was modest (M = 300, from 400 Gibbs iterations with 100 discarded as warm-up), which is too few to locate the 2.5th and 97.5th percentiles precisely and is therefore itself a source of interval instability.

We call the result a credible interval, not a confidence interval, because it is a percentile range of a distribution for the quantity being estimated, not the output of a procedure constructed to cover the true value in 95% of repeated samples. No sampling distribution is inverted, so the interval carries no guarantee of frequentist coverage. The distinction is not terminological: in the simulation, where BAF_true is known by construction, the interval contained it in `r pct(100 * ciw$coverage_overall, 1)` of repetitions, below the nominal 95%, and coverage was lower among narrower intervals (`r pct(100 * ciw$coverage_narrow, 1)`) than among wider ones (`r pct(100 * ciw$coverage_wide, 1)`). One reason is visible in the construction itself: the interval omits the sampling variability of the effect estimate, a second source of error in BAF̂. Reported as a credible interval with its coverage stated, this is an honest statement that the propagated distribution is too narrow rather than a failure of the method.

Each interval is summarised by its half-width, (BAF_hi − BAF_lo)/2, and not by its two bounds. BAF is bounded, so the bounds are truncated at 0 and 1 and carry information about location as well as precision; the half-width extracts the precision component in BAF units, on the same bounded scale as the estimate itself, which is what lets the lookup table of Step 2 classify an estimate jointly by where it falls on the BAF scale and by how wide its interval is, splitting the half-width at its median across the simulation (`r fmt_num(ciw$median_half_width, 3)`) and returning the higher of the two stratum-specific probabilities. The half-width is not a standard error: because the bounds are truncated, it is half of the [0, 1] scale that the interval occupies, and near either end of the scale it is attenuated by that truncation.

### S1.5 Justification of the simulation factor ranges

Factor levels were not chosen to represent a survey of cardiovascular observational research. Each level was anchored either to a quantity measured in the present case study or to a published estimate of adherence-related bias, as follows.

```{r sup-rr-vars, include=FALSE}
rr_psi_sup <- paste(rr_of(max(unlist(des$psi_levels))),  "to",
                    rr_of(min(unlist(des$psi_levels))))   # -0.01 -> -0.40
rr_mu_sup  <- paste(rr_of(max(unlist(des$mu_b_levels))),  "to",
                    rr_of(min(unlist(des$mu_b_levels))))   # -0.05 -> -0.40
rr_exv_sup <- rr_of(0.15)                                  # 偏移量，未进 JSON
```

| Factor | Levels | Values | Justification | Source |
|---|---|---|---|---|
| True effect, ψ | `r length(des$psi_levels)` | `r fmt_signed(max(unlist(des$psi_levels)), 2)` to `r fmt_signed(min(unlist(des$psi_levels)), 2)` | Spans marginal to large protective effects on the log relative-risk scale (RR, `r rr_psi_sup`) and brackets the beta-blocker estimate observed in the case study (ln RR, `r fmt_signed(cases[[1]]$log_rr_uncal, 2)`) | Present case study |
| Bias center, μB | `r length(des$mu_b_levels)` | `r fmt_signed(max(unlist(des$mu_b_levels)), 2)` to `r fmt_signed(min(unlist(des$mu_b_levels)), 2)` | (RR, `r rr_mu_sup`). Approaches, without reaching, the healthy-adherer effect reported for adherence-related exposures (odds ratio, 0.56; ln, −0.58) and brackets the empirical null fitted to the case-study panel (μ̂B, `r fmt_signed(cases[[1]]$mu_b, 2)`) | Simpson 2006; present case study |
| Standard error of the target estimate, σψ | `r length(des$sigma_ps_levels)` | `r fmt_num(des$sigma_ps_levels[1], 2)`, `r fmt_num(des$sigma_ps_levels[2], 2)` | Closely matches the standard errors of the two case-study estimates (`r fmt_num(se_obs_of(cases[[1]]), 3)` and `r fmt_num(se_obs_of(cases[[2]]), 3)`), the smaller falling marginally below the lower grid level | Present case study |
| Number of negative controls, K | `r length(des$K_levels)` | `r paste(des$K_levels, collapse = ", ")` | 12 is the number of controls in the cardiovascular panel available in MIMIC-IV; 25 is a moderately larger panel; 50 tests whether estimator performance keeps improving or plateaus once the panel far exceeds the size typically assembled in practice | Present case study; Schuemie 2018 |
| Exchangeability violation | `r length(des$exV_levels)` | `r des$exV_levels[1]`, `r des$exV_levels[2]` | 0 corresponds to an assumption that holds; 0.3 shifts 30% of controls by +0.15 (RR, `r rr_exv_sup`), a deliberately arbitrary perturbation chosen to break the assumption rather than to reproduce any empirical estimate | Design choice |

Three respects in which the grid does not bracket the case study should be read as limitations of the simulation, all of which make it more favorable to the estimator than the case study itself. First, the between-control standard deviation was held fixed at 0.05, below the `r fmt_num(cases[[1]]$sigma_b, 3)` fitted to the case-study panel. Second, control standard errors were drawn from a uniform distribution on 0.03 to 0.12, whereas the observed controls ranged from 0.025 to 0.302 with 3 controls above 0.24. Third, the guideline-directed medical therapy estimate (ln RR, `r fmt_signed(cases[[2]]$log_rr_uncal, 2)`) lies between grid levels, within the ψ grid whose most extreme value is `r fmt_signed(min(unlist(des$psi_levels)), 2)`.

---

## S2. Part I supplementary: Bland-Altman agreement

The interior analysis (primary) excludes the ψ = -0.01 boundary level to avoid near-boundary pathologies. The full analysis includes it. Both are reported below.

```{r ba-table}
ba <- data.frame(
  Analysis = c("Interior (primary)", "Full (incl. boundary)"),
  N = c(p1_int$n, p1_full$n),
  `Mean diff (BAFhat-BAFtrue)` = c(p1_int$bias, p1_full$bias),
  `SD diff` = c(p1_int$sd_diff, p1_full$sd_diff),
  `LoA low` = c(p1_int$loa_lo, p1_full$loa_lo),
  `LoA high` = c(p1_int$loa_hi, p1_full$loa_hi),
  `CCC` = c(p1_int$ccc, p1_full$ccc),
  `Prop. bias slope` = c(p1_int$slope, p1_full$slope),
  MAE = c(p1_int$mae, p1_full$mae),
  RMSE = c(p1_int$rmse, p1_full$rmse)
)
kable(ba, digits = c(0, 0, 3, 3, 3, 3, 3, 4, 3, 3),
      caption = "Table S1. Bland-Altman agreement of BAFhat as an estimator of BAFtrue.")
```

The worst single-condition mean signed difference among the `r fmt_num(NUM$part1$n_conditions_interior, 0)` interior conditions was `r fmt_num(NUM$part1$interior_worst_cond_bias, 3)`. The proportional-bias slope regression rejected the null of no proportional bias in both analyses (interior slope `r fmt_num(p1_int$slope, 4)`, 95% CI `r fmt_num(p1_int$slope_lo, 4)` to `r fmt_num(p1_int$slope_hi, 4)`; full slope `r fmt_num(p1_full$slope, 4)`, 95% CI `r fmt_num(p1_full$slope_lo, 4)` to `r fmt_num(p1_full$slope_hi, 4)`). The estimator systematically underestimates BAF_{true}, and the underestimation grows with the magnitude of BAF_{true}.

---

## S3. Part II supplementary: decision-rule detail

### S3.1 Full 12-bucket lookup table

Each bucket reports the count of repetitions with narrow and wide credible intervals, the probability of being bias-dominated under the narrow reading, the wide reading, and the conservative max(narrow, wide) reading, together with the verdict zone.

```{r lookup-table}
lk_rows <- function(L) {
  data.frame(
    Bin = sapply(L, `[[`, "bin"),
    `BAF centre` = sapply(L, `[[`, "bf_centre"),
    `n narrow` = sapply(L, `[[`, "n_narrow"),
    `n wide` = sapply(L, `[[`, "n_wide"),
    `pBD narrow` = sapply(L, function(x) kpct(x$p_bd_narrow, 2)),
    `pBD wide` = sapply(L, function(x) kpct(x$p_bd_wide, 2)),
    `pBD cons.` = sapply(L, function(x) kpct(x$p_bd_conservative, 2)),
    Verdict = sapply(L, `[[`, "verdict")
  )
}
kable(lk_rows(lookup), digits = c(0, 3, 0, 0, 0, 0, 0, 0),
      caption = "Table S2. Twelve-bucket equal-frequency lookup table mapping BAFhat x interval width to probability of being bias-dominated.")
```

### S3.2 Three readings of the lookup

The lookup supports a pooled reading (CI width ignored), a width-specific reading, and the conservative max(narrow, wide) reading used in the manuscript.

```{r reading-table}
rr <- cirole$rules
rd <- data.frame(
  Reading = sapply(rr, `[[`, "reading"),
  `Misuse rate (%)` = sapply(rr, function(x) fmt_pct(x$bd_among_declared_pct, 2)),
  `Declare usable (%)` = sapply(rr, function(x) fmt_pct(x$declare_usable_pct, 1)),
  # spec_bd_pct 是 R/18 的历史命名，它算的是 P(扣下 | 真偏倚主导)，即 BD 视角的
  # 灵敏度而非特异度；表头按正确语义写，变量名保持不变（改变量名要动 R/18）。
  `Sens ED (%)` = sapply(rr, function(x) fmt_pct(x$sens_ed_pct, 1)),
  `Sens BD (%)` = sapply(rr, function(x) fmt_pct(x$spec_bd_pct, 1))
)
kable(rd, digits = 0,
      caption = paste0("Table S3. Comparison of three readings of the lookup table. ",
                       "Sens ED, share of truly effect-dominated repetitions declared usable; ",
                       "Sens BD, share of truly bias-dominated repetitions withheld."))
```

The conservative reading changed `r pct(cirole$pct_verdict_changed_spec_vs_pool, 2)` of verdicts relative to the pooled reading and `r pct(cirole$pct_verdict_changed_cons_vs_pool, 2)` relative to the width-specific reading; the maximum absolute within-bin probability gap was `r fmt_num(cirole$max_abs_prob_gap_within_bin, 3)`, and adding CI half-width moved the AUC by only `r fmt_num(cirole$delta_auc_from_ci_width, 5)`.

### S3.3 Nested strategy comparison

Strategies R0 through R4 add one component at a time; C1 and C2 are stand-alone comparators that do not use the calibrated P-value. Every rule is scored twice, once from each viewpoint of Table 4 in the main text, so Table S4 reports eight operating characteristics per rule. The two readings are not mirror images because truth has three levels and the mixed zone is negative in both.

```{r strategy-table}
RULE_LABEL <- c(
  R0 = "R0 uncalibrated P < .05 (no bias correction)",
  R1 = "R1 calibrated P < .05 (layer 1 only)",
  R2 = "R2 layer 1, plus BAF point estimate below 0.5, the bias-dominated threshold",
  R3 = "R3 layer 1, plus the whole credible interval of BAF below 0.5, the bias-dominated threshold",
  R4 = "R4 layer 1, plus probability of bias dominance below 15%",
  C1 = "C1 BAF point estimate below 0.5, the bias-dominated threshold, no layer 1",
  C2 = "C2 the whole credible interval of BAF below 0.5, the bias-dominated threshold, no layer 1"
)
# 双视角 10 列。旧版本的 "Spec BD" 列命名有误：它拿到的是
# strat$spec_bd_pct，其实是 BD 视角的 sensitivity（真偏倚主导且被扣下），
# 不是 specificity。R/23 重命名后数字不变、语义对齐。
# Misuse 列删掉：它等于 100 减 NPV(BD)，保留会重复且占宽度。
stopifnot(identical(dual$strategies$rule,
                    sub(" .*", "", sapply(strat, `[[`, "strategy"))))
st <- data.frame(
  # 规则标签：管线里的内部说法换成读者能独立读懂的表述
  # （与正文 Table 4 的 Definition 列同口径，避免两处说法打架）
  Strategy            = unname(RULE_LABEL[dual$strategies$rule]),
  `Declare usable (%)` = fmt_pct(dual$strategies$declare_usable_pct, 1),
  `Sens, ED (%)`      = fmt_pct(dual$strategies$sens_ed, 1),
  `Spec, ED (%)`      = fmt_pct(dual$strategies$spec_ed, 1),
  `PPV, ED (%)`       = fmt_pct(dual$strategies$ppv_ed, 1),
  `NPV, ED (%)`       = fmt_pct(dual$strategies$npv_ed, 1),
  `Sens, BD (%)`      = fmt_pct(dual$strategies$sens_bd, 1),
  `Spec, BD (%)`      = fmt_pct(dual$strategies$spec_bd, 1),
  `PPV, BD (%)`       = fmt_pct(dual$strategies$ppv_bd, 1),
  `NPV, BD (%)`       = fmt_pct(dual$strategies$npv_bd, 1),
  check.names = FALSE
)
kable(st, digits = 0, align = c("l", rep("r", 9)),
      caption = paste0(
        "Table S4. Nested strategy comparison, read from two viewpoints. ",
        "In the effect-dominated view (ED) a positive call is to declare a repetition usable and a positive truth is to be truly effect-dominated; ",
        "in the bias-dominated view (BD) a positive call is to withhold and a positive truth is to be truly bias-dominated. ",
        "Sens, share of truly positive repetitions the rule called positive; Spec, share of truly negative repetitions the rule called negative; ",
        "PPV, share of repetitions the rule called positive that were truly positive; NPV, share the rule called negative that were truly negative. ",
        "The misuse rate reported in Table 4 of the main text is 100 minus NPV in the BD view. Truth has three levels (effect-dominated ",
        kpct(dual$truth_share$effect_dominated, 1), ", mixed ", kpct(dual$truth_share$mixed, 1),
        ", bias-dominated ", kpct(dual$truth_share$bias_dominated, 1),
        "), and the mixed zone is negative in both views, so the views are not mirror images and no pair of columns sums to 100."))
```

### S3.4 Decision flow by stage

```{r flow-table}
fl <- data.frame(
  Stage = sapply(flow, `[[`, "stage"),
  N = sapply(flow, `[[`, "n"),
  `Bias-dominated (n)` = sapply(flow, `[[`, "bd_true_n"),
  `BD share (%)` = sapply(flow, function(x) fmt_pct(x$bd_share_pct, 1))
)
kable(fl, digits = 0,
      caption = "Table S5. Decision flow across layers of the rule.")
```

### S3.5 Verdict versus truth (3 by 3 confusion)

```{r confusion-table}
cf <- data.frame(
  Truth = sapply(confusion3x3, `[[`, "truth"),
  `bias-dominated` = sapply(confusion3x3, `[[`, "bias-dominated"),
  `mixed` = sapply(confusion3x3, `[[`, "mixed"),
  `effect-dominated` = sapply(confusion3x3, `[[`, "effect-dominated"),
  `Row total` = sapply(confusion3x3, `[[`, "row_total")
)
kable(cf, digits = 0,
      caption = "Table S6. Confusion matrix of assigned verdict zone against simulated truth.")
```

Zone accuracy was `r pct(NUM$part2$zone_accuracy$overall, 1)` overall, `r pct(NUM$part2$zone_accuracy$bias_dominated, 1)` among bias-dominated, `r pct(NUM$part2$zone_accuracy$mixed, 1)` among mixed, and `r pct(NUM$part2$zone_accuracy$effect_dominated, 1)` among effect-dominated repetitions.

### S3.6 Subgroup robustness

```{r subgroup-table}
sg <- data.frame(
  Subgroup = sapply(subg, `[[`, "subgroup"),
  N = sapply(subg, `[[`, "n"),
  `Layer1 pass (%)` = sapply(subg, function(x) fmt_pct(x$layer1_pass_pct, 1)),
  `BD among pass (%)` = sapply(subg, function(x) fmt_pct(x$bd_among_layer1_pct, 1)),
  `Declare usable (%)` = sapply(subg, function(x) fmt_pct(x$declare_usable_pct, 1)),
  `BD among usable (%)` = sapply(subg, function(x) fmt_pct(x$bd_among_declared_pct, 2)),
  `Median half-width` = sapply(subg, `[[`, "median_half_width")
)
kable(sg, digits = c(0, 0, 1, 1, 1, 2, 3),
      caption = paste0("Table S7. Misuse rate across ", length(subg), " pre-specified subgroups."))
```

### S3.7 Discrimination

```{r discrimination-table}
dc <- data.frame(
  Predictor = sapply(disc, `[[`, "predictor"),
  `AUC` = sapply(disc, function(x) fmt_num(x$auc, 3)),
  `LR chi2 note` = sapply(disc, function(x) ifelse(x$note == "", "-", x$note))
)
kable(dc, digits = 0,
      caption = "Table S8. Discrimination (AUC). The area under the curve is read from the bias-dominated view, meaning that withholding a repetition is a positive call and being truly bias-dominated is a positive truth. Figure 8 of the main text reports the same three scores from both viewpoints.")
```

The likelihood-ratio chi-square test for adding credible-interval half-width to BAF̂ was χ² = `r sub(".*chi2 = ([0-9.]+).*", "\\1", disc[[3]]$note)`, df = 1, p < .001. The odds ratio per unit increase in CI half-width was `r fmt_num(NUM$part2$or_half_width, 2)`.

### S3.8 Credible-interval width and coverage

The median half-width was `r fmt_num(ciw$median_half_width, 3)` (quartiles `r fmt_num(ciw$quartiles_half[1], 3)`, `r fmt_num(ciw$quartiles_half[3], 3)`). Coverage of the Gibbs posterior 95% credible interval was `r pct(100 * ciw$coverage_overall, 1)` overall, `r pct(100 * ciw$coverage_narrow, 1)` for narrow intervals, and `r pct(100 * ciw$coverage_wide, 1)` for wide intervals. Mean half-width by K was `r fmt_num(ciw$mean_half_by_K[1], 3)` for K = 12 and `r fmt_num(ciw$mean_half_by_K[2], 3)` for K = 25. The calibrated P-value at the 0.05 threshold declared `r pct(100 * l1$pass_rate_overall, 1)` of repetitions passable versus `r pct(100 * l1$naive_pass_rate, 1)` for the uncalibrated P-value; the pass rate among bias-dominated truth was `r pct(100 * l1$pass_rate_by_truth$bias_dominated, 1)`, among mixed `r pct(100 * l1$pass_rate_by_truth$mixed, 1)`, and among effect-dominated `r pct(100 * l1$pass_rate_by_truth$effect_dominated, 1)`.

---

## S4. Part III supplementary: case-study estimation

### S4.1 Negative-control empirical null

The jointly estimated empirical null had center μ̂_B = `r fmt_num(cases[[1]]$mu_b, 3)` and scale σ̂_B = `r fmt_num(cases[[1]]$sigma_b, 3)`; the Shapiro-Wilk test of the `r fmt_num(nc$K, 0)` negative-control residuals gave p = `r fmt_num(nc$shapiro_p, 3)`, and the calibrated log relative risks of the controls ranged from `r fmt_num(nc$logrr_min, 3)` to `r fmt_num(nc$logrr_max, 3)`.

### S4.2 Full case estimation

```{r case-table}
cs <- data.frame(
  Question = c(cases[[1]]$exposure, cases[[2]]$exposure),
  `Uncal RR` = c(cases[[1]]$rr_uncal, cases[[2]]$rr_uncal),
  `Cal RR` = c(cases[[1]]$rr_cal, cases[[2]]$rr_cal),
  `Cal P` = c(cases[[1]]$cal_p, cases[[2]]$cal_p),
  `BAFhat` = c(cases[[1]]$bf, cases[[2]]$bf),
  `BAF lo` = c(cases[[1]]$bf_lo, cases[[2]]$bf_lo),
  `BAF hi` = c(cases[[1]]$bf_hi, cases[[2]]$bf_hi),
  `p bias-dom` = c(cases[[1]]$p_bias_dominated, cases[[2]]$p_bias_dominated),
  Verdict = c(cases[[1]]$verdict, cases[[2]]$verdict)
)
kable(cs, digits = c(0, 3, 3, 4, 3, 3, 3, 3, 0),
      caption = "Table S9. Full estimation output for the two clinical questions.")
```

### S4.3 Sensitivity analyses (question 1)

Leave-one-out across the `r fmt_num(nc$K, 0)` negative controls gave BAF̂ ranging from `r fmt_num(sens$loo_bf_min, 3)` to `r fmt_num(sens$loo_bf_max, 3)` (fall-back BAF̂ = `r fmt_num(sens$loo_fall_bf, 3)`). Replacement of the Super Learner library gave BAF̂ = `r fmt_num(sens$sl_bf[1], 3)` (`r fmt_num(sens$sl_bf_lo[1], 3)` to `r fmt_num(sens$sl_bf_hi[1], 3)`), `r fmt_num(sens$sl_bf[2], 3)` (`r fmt_num(sens$sl_bf_lo[2], 3)` to `r fmt_num(sens$sl_bf_hi[2], 3)`), and `r fmt_num(sens$sl_bf[3], 3)` (`r fmt_num(sens$sl_bf_lo[3], 3)` to `r fmt_num(sens$sl_bf_hi[3], 3)`) under the three alternative libraries. The Gaussian-process propensity-score perturbation gave calibrated relative risk `r fmt_num(sens$gp_cal_rr[1], 3)` to `r fmt_num(sens$gp_cal_rr[2], 3)` and BAF̂ `r fmt_num(sens$gp_bf[1], 3)` to `r fmt_num(sens$gp_bf[2], 3)`. All sensitivity analyses preserved the "not usable as effect evidence" verdict for question 1 and the "competitive, no verdict" zone for question 2.

---

## S5. Additional sensitivity and validation analyses

### S5.1 Held-out validation of the lookup table

The lookup is built from all `r format(des$n_total, big.mark = ",")` repetitions and then used to score the reporting rules on the same repetitions, so its reported operating characteristics are in-sample. We bounded the resulting optimism by rebuilding the entire lookup, including the bin edges, the median half-width that separates narrow from wide intervals, and the per-bin probabilities, on training data alone, and then applying it unchanged to held-out data. Two schemes were used, each repeated across five folds. In the first, 20% of repetitions were held out within every condition, so that both training and test data span the whole factorial grid; the misuse rate of every rule was then unchanged to two decimal places, which supports the argument that at least `r format(des$n_reps_per_cond, big.mark = ",")` repetitions per cell makes the in-sample optimism negligible when only repetitions are swapped. In the second, 20% of the `r fmt_num(des$n_conditions, 0)` conditions were held out entirely, which additionally tests grid cells that never contributed to the lookup; the share declared usable was essentially unchanged, and the misuse rate of the strictest rule (R4) rose from 2.2% to 2.4%, the largest excursion across the five folds being 4.1% and the largest increase for any rule 0.23 percentage points (Table S10). Generalisation to unseen cells therefore costs little, but it is not free, and the ordering of the rules is preserved.

```{r sup-holdout}
ho <- read.csv(file.path(TAB_DIR, "v39_lookup_holdout.csv"), stringsAsFactors = FALSE)
ho$rule_short <- sub(" .*", "", ho$rule)
ho$scheme_lab <- ifelse(ho$scheme == "rep", "20% of repetitions held out", "20% of conditions held out")
ho <- ho[order(match(ho$scheme, c("rep", "condition")),
               match(ho$rule_short, c("R0","R1","R2","R3","R4","C1","C2"))), ]
ht <- data.frame(
  `Held out` = ho$scheme_lab,
  Rule = ho$rule_short,
  `Declared usable, training (%)` = kpct(ho$declare_train, 1),
  `Declared usable, test (%)` = kpct(ho$declare_test, 1),
  `Misuse, training (%)` = kpct(ho$misuse_train, 2),
  `Misuse, test (%)` = kpct(ho$misuse_test, 2),
  `Misuse across folds (%)` = sprintf("%.1f\u2013%.1f", ho$misuse_test_min, ho$misuse_test_max),
  check.names = FALSE, stringsAsFactors = FALSE
)
kable(ht, caption = paste0(
  "Table S10. Five-fold held-out validation of the BAF reliability lookup, for the seven reporting ",
  "rules of Table 4. The lookup, including its bin edges, its median half-width, and its per-bin ",
  "probabilities, was rebuilt on training data only and applied unchanged to the held-out data; the ",
  "training columns report its performance on the data used to build it, so the gap between the ",
  "training and test columns is the in-sample optimism. Misuse is the share of repetitions declared ",
  "usable that were in truth bias-dominated."))
```

### S5.2 Joint propagation of empirical-null and target-estimate uncertainty

The reported credible interval propagates draws of the empirical-null mean while holding the observed target estimate fixed, so it omits the sampling variability of that estimate and undercovers. To quantify what this omission costs, we recomputed the interval with the omitted term added, drawing the calibrated target estimate from a normal distribution centred on its observed value with its own standard error. Both intervals were computed within the same repetitions, so the comparison is paired. Across the settings examined, coverage rose from 78.8% to 97.7%, the median half-width widened from 0.132 to 0.222, and the yield of the strictest interval rule fell from 24.3% to 13.2% of repetitions, with 11.2% of repetitions changing that rule's decision (Table S11). The undercoverage reported in the main text is therefore attributable specifically to the omitted term, and the yield of the rules that consume the interval endpoints is a conservative lower bound.

```{r sup-joint}
ij <- read.csv(file.path(TAB_DIR, "v39_interval_joint_propagation.csv"), stringsAsFactors = FALSE)
ij2 <- data.frame(
  Scope = ij$scope,
  N = format(ij$n, big.mark = ","),
  `Coverage, reported (%)` = kpct(ij$coverage_current, 1),
  `Coverage, joint (%)` = kpct(ij$coverage_joint, 1),
  `Median half-width, reported` = fmt_num(ij$median_halfwidth_current, 3),
  `Median half-width, joint` = fmt_num(ij$median_halfwidth_joint, 3),
  `Rule yield, reported (%)` = kpct(ij$R3_yield_current, 1),
  `Rule yield, joint (%)` = kpct(ij$R3_yield_joint, 1),
  `Rule misuse, reported (%)` = kpct(ij$R3_misuse_current, 2),
  `Rule misuse, joint (%)` = kpct(ij$R3_misuse_joint, 2),
  check.names = FALSE, stringsAsFactors = FALSE
)
kable(ij2, caption = paste0(
  "Table S11. Interval coverage, precision, and rule performance when the sampling variability of ",
  "the calibrated target estimate is propagated jointly with that of the empirical null, compared ",
  "with the reported interval, which holds the target estimate fixed. Both intervals were computed ",
  "within the same ", format(ij$n[1], big.mark = ","), " repetitions, so the comparison is paired. ",
  "Rule is the strictest interval rule (R3). The joint interval is wider and attains nominal ",
  "coverage, at the cost of a lower yield."))
```

### S5.3 Missing-data handling for the guideline-directed medical therapy exposure

Of the 14,677 patients in the question-2 cohort, 3,494 (23.8%) had an indeterminate guideline-directed medical therapy score at baseline. The primary analysis classified them as not optimized, which is a design decision rather than a data fact. We therefore repeated the question-2 estimation under four conventions: indeterminate as suboptimal (the published analysis), complete case, indeterminate as optimized, and multiple imputation, in which the exposure was drawn from a logistic model of exposure on the ten baseline covariates and combined across ten imputations by Rubin's rules. The target estimate moved substantially across conventions (uncalibrated relative risk 0.68 to 0.91), but the calibrated P-value never fell below .05 and no convention produced a usable determination; the bias fraction rose from 0.51 under the published convention to between 0.51 and 0.66 under the others, and three of the four conventions moved the determination to not usable as effect evidence rather than the competitive determination returned by the published convention (Table S12). The published convention is thus the most favourable of the four, so the qualitative conclusion is not an artefact of it. The negative-control panel was fitted with the exposure contrast that counts all beta-blocker agents, whereas the target estimates use the guideline-restricted contrast; refitting the panel to match widens the empirical null (SD `r fmt_num(nc_matched_sigma,3)` versus `r fmt_num(sup_sigma_b,3)`) and raises the question-2 calibrated P value above .05 (`r fmt_num(nc_matched_calp_q2,3)` versus `r fmt_num(sup_calp_q2,3)`), so the question-2 layer-1 pass is not robust to this choice.

```{r sup-missing}
gms <- read.csv(file.path(TAB_DIR, "v39_gdmt_missing_sensitivity.csv"), stringsAsFactors = FALSE)
gm <- data.frame(
  `Indeterminate exposure handled as` = gms$tag,
  N = format(gms$n, big.mark = ","),
  `Uncalibrated RR` = fmt_num(gms$rr, 3),
  `Uncalibrated RR 95% CI` = sprintf("%.3f\u2013%.3f", gms$rr_lo, gms$rr_hi),
  `Calibrated P` = sub("^0\\.", ".", fmt_num(gms$cal_p, 4)),
  BAF = fmt_num(gms$bf, 3),
  `BAF 95% credible interval` = sprintf("%.3f\u2013%.3f", gms$bf_lo, gms$bf_hi),
  `P(bias-dominated)` = fmt_num(gms$p_bias_dominated, 3),
  Determination = gms$verdict,
  check.names = FALSE, stringsAsFactors = FALSE
)
kable(gm, caption = paste0(
  "Table S12. Question 2 under four conventions for the ", format(gms$n[1] - gms$n[2], big.mark = ","),
  " patients (", kpct(100 * (gms$n[1] - gms$n[2]) / gms$n[1], 1), ") with an indeterminate ",
  "guideline-directed medical therapy score. The bias fraction and its credible interval are computed ",
  "with the same bootstrap estimator and n_boot used for the published analysis, and the ",
  "determination is read from the same lookup table, so the only quantity that varies across rows is ",
  "the target estimate. The first row reproduces the published question-2 result exactly."))
```

---

## S6. Software and reproducibility


The estimator is implemented in the R package EmpiricalCalibration [@schuemie2018empirical]; the BAF extension is provided by the accompanying R code archived with this manuscript. The analysis pipeline reads MIMIC IV and MIMIC IV ECG under approved data-use agreements; the derived negative-control and estimation tables, the simulation frame, and the single JSON of reported numbers are released with the manuscript repository. The companion interactive explainer consumes the same JSON so that every figure and table in the manuscript and the explainer are guaranteed to agree.
