---
title: "Quantifying Systematic Bias in Observational Causal Estimates Using Negative Controls: The Bias Fraction"
author:
  - name: Zeng Zhechun
    degrees: MD, MPH
    affiliation: "1,2"
    email: zengzhechun@icloud.com
    corresponding: true
  - name: Wang Jinwen
    degrees: MD, PhD
    affiliation: "1,2"
  - name: Zuo Huijuan
    degrees: MD, MPH
    affiliation: "1,2"
  - name: Shu Lixia
    degrees: PhD
    affiliation: "1,2"
date: "2026-08-27"
bibliography: references.bib
csl: american-medical-association.csl
format:
  docx:
    toc: false
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

^1^ Beijing Anzhen Hospital, Capital Medical University, Beijing, China

^2^ Beijing Institute of Heart, Lung and Blood Vessel Diseases, Beijing, China

**Corresponding author:** Zeng Zhechun, MD, MPH
Beijing Anzhen Hospital, Capital Medical University, Tongzhou Campus, No. 225 Songzhuang South 1st Street, Tongzhou District, Beijing 101149, China
Email: zengzhechun@icloud.com

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(jsonlite)
  library(ggplot2)
})
source("../R/00_config.R")
source("../R/01_bsr_core.R")
library(EmpiricalCalibration)

# Single source of truth for every number quoted in the manuscript
NUM <- jsonlite::fromJSON(file.path(TAB_DIR, "v38_all_numbers.json"),
                          simplifyVector = FALSE)

# Dual-view operating characteristics: the same rules R0-R4, C1-C2 scored twice,
# once taking "declared usable" as the positive call (effect-dominated view, ED)
# and once taking "withheld" as the positive call (bias-dominated view, BD), each
# with sensitivity, specificity, PPV and NPV. Written by R/23_dual_view_metrics.R.
dual <- jsonlite::fromJSON(file.path(TAB_DIR, "v38_dual_view.json"),
                           simplifyVector = TRUE)

# Convenience accessors
des       <- NUM$design
p1_int    <- NUM$part1$interior
p1_full   <- NUM$part1$full
l1        <- NUM$part2$layer1
ciw       <- NUM$part2$ci_width
disc      <- NUM$part2$discrimination
strat     <- NUM$part2$strategies
lookup    <- NUM$part2$lookup
flow      <- NUM$part2$flow
subg      <- NUM$part2$subgroups
cirole    <- NUM$part2$ci_width_role
cases     <- NUM$part3$cases
sens      <- NUM$part3$sensitivity

fmt_pct <- function(x, d = 1) sprintf(paste0("%.", d, "f"), as.numeric(x))
fmt_num <- function(x, d = 2) sprintf(paste0("%.", d, "f"), as.numeric(x))
# fmt_pct formats as-is; kpct is for fields stored as proportions (eg
# des$truth_share = 0.525) that must be shown as 52.5%.
kpct <- function(x, d = 1) paste0(fmt_pct(100 * as.numeric(x), d), "%")

BB  <- cases[[1]]
GDM <- cases[[2]]
```

## Abstract

**Importance.** Observational studies dominate evidence in cardiovascular outcomes research, yet no widely adopted reporting standard requires authors to disclose *how much* of an observed association is attributable to systematic error. The negative-control calibration framework of the Observational Health Data Sciences and Informatics consortium returns a binary answer: a calibrated P-value either crosses 0.05 or it does not. That answer conflates "is there a signal" with "is the signal credible."

**Objective.** To define the Bias Fraction (BF), a continuous estimator in [0, 1] that quantifies the share of an observed log relative risk that is attributable to systematic bias, together with its 95% credible interval, and to evaluate whether reporting calibrated P-value plus BF plus interval width produces more defensible decisions than reporting the calibrated P-value alone.

**Design.** Three-part study: (i) metric definition and interval construction; (ii) Monte Carlo simulation across `r fmt_num(des$n_conditions,0)` conditions × `r fmt_num(des$n_reps_per_cond,0)` repetitions = `r format(des$n_total, big.mark = ",")` simulated target estimates, with truth known by construction; (iii) two-arm beta-blocker target trial emulation in MIMIC IV with targeted maximum likelihood estimation (TMLE) and `r fmt_num(NUM$part3$nc$K,0)` negative-control outcomes.

**Setting.** MIMIC IV (critical care) and MIMIC IV ECG, 2008 to 2019.

**Participants.** Adults with a first intensive care unit stay and a recorded 12-lead ECG.

**Exposures.** Two clinical questions are evaluated: (1) reaching at least 50% of guideline-recommended target dose of an index beta-blocker within 7 days of admission, sustained to 12 months, versus lower or no exposure; (2) receiving guideline-directed medical therapy (GDMT) with at least two of three foundational drug classes at adequate dose for 12 months versus fewer.

**Main Outcomes and Measures.** One-year all-cause mortality. Primary analytic outputs: (i) the calibrated P-value from a negative-control empirical null distribution with `r fmt_num(NUM$part3$nc$K,0)` outcomes; (ii) BF̂ with a Markov chain Monte Carlo 95% credible interval; (iii) a verdict drawn from a truth-calibrated lookup table mapping the BF̂ × interval-width cell to a probability of being bias-dominated.

**Results.** Across `r format(des$n_total, big.mark = ",")` simulated repetitions, the calibrated P-value alone yielded a misuse rate (declared usable while truth was bias-dominated) of `r fmt_pct(strat[[2]]$bd_among_declared_pct,2)`. Adding the BF point estimate reduced misuse to `r fmt_pct(strat[[3]]$bd_among_declared_pct,2)`, and requiring the whole credible interval of BF to lie below 1/3 reduced misuse to `r fmt_pct(strat[[4]]$bd_among_declared_pct,2)`. Requiring in addition a calibrated probability of bias dominance below 15% reduced misuse to `r fmt_pct(strat[[5]]$bd_among_declared_pct,2)` at the cost of declaring `r fmt_pct(strat[[5]]$declare_usable_pct,1)` of estimates usable (versus `r fmt_pct(strat[[2]]$declare_usable_pct,1)` for the calibrated P-value alone). Read from two viewpoints, the BF estimator achieved an area under the receiver operating characteristic curve of `r fmt_num(disc[[2]]$auc, 3)` (bias-dominated view) and `r fmt_num(dual$auc$auc_ed[2], 3)` (effect-dominated view), compared with `r fmt_num(disc[[1]]$auc, 3)` and `r fmt_num(dual$auc$auc_ed[1], 3)` for the calibrated P-value. In MIMIC IV, question 1 produced a calibrated P-value of `r fmt_num(BB$cal_p, 3)` and BF̂ = `r fmt_num(BB$bf, 3)` (95% credible interval `r fmt_num(BB$bf_lo, 3)` to `r fmt_num(BB$bf_hi, 3)`); the apparent one-year mortality reduction of `r fmt_pct(1 - BB$rr_uncal, 1)` was entirely attributable to systematic bias and the association was not usable as effect evidence. Question 2 produced a calibrated P-value of `r fmt_num(GDM$cal_p, 4)`, passing layer 1, but BF̂ = `r fmt_num(GDM$bf, 3)` placed it in the "competitive, no verdict" zone.

**Conclusions and Relevance.** Pairing the calibrated P-value with BF and its 95% credible interval turns a binary credibility judgement into a three-channel report (signal existence, effect size, and bias share). Under the most rigorous design attainable in observational research, both clinical questions in this study failed to qualify as effect evidence, with question 1 demonstrating complete bias domination. The reporting standard proposed here can be applied to any observational study that includes negative controls, without requiring additional data.

---

## Introduction

Observational studies are the dominant source of evidence for cardiovascular effectiveness and safety questions that cannot be answered by randomized trials. They are also vulnerable to systematic error that survives the most rigorous design choices available to observational research: explicit eligibility criteria, time-zero alignment, intention-to-treat and per-protocol analyses, and modern estimators such as longitudinal targeted maximum likelihood estimation (LTMLE) that handle time-varying confounding and informative censoring [@lendle2017ltmle; @vanderlaan2011targeted]. A common reaction to residual error is to require authors to perform a sensitivity analysis. Sensitivity analyses ask the analyst to stipulate how much unmeasured confounding would have to exist to overturn the result, then return a qualitative judgement. They do not quantify how much of the observed association is plausibly attributable to bias.

The Observational Health Data Sciences and Informatics (OHDSI) consortium has proposed a complementary practice: estimate an empirical null distribution from a panel of negative-control outcomes, jointly estimate its center and scale, and use that distribution to produce a calibrated P-value and a calibrated effect estimate [@schuemie2018empirical]. The calibrated P-value separates the question "does a signal exist after removing systematic error" from the question "how large is the effect." That is a useful step, but it returns a binary answer for the first question and a point estimate for the second, leaving the analyst to decide, without guidance, whether the residual estimate is dominated by bias or by effect.

This manuscript develops a third quantity, the Bias Fraction (BF), defined on the log relative risk scale as the share of the absolute observed association attributable to the absolute estimated systematic bias:

$$
BF = \frac{|\hat{\mu}_B|}{|\hat{\mu}_B| + |\tilde{\psi}|},
$$

where $\hat{\mu}_B$ is the estimated center of the empirical null and $\tilde{\psi}$ is the calibrated log relative risk. BF lies in [0, 1], equals 0 when the calibrated effect absorbs the entire magnitude, and equals 1 when systematic bias absorbs the entire magnitude. We use BF as the primary metric and BF̂ and its 95% credible interval as the principal object of inference. BF is related to the bias-to-effect ratio by $BF = BER / (1 + BER)$; BER is reported as an auxiliary summary to maintain consistency with prior reports from this program of work.

The manuscript is organised in three sequential parts. Part I defines BF and the credible-interval construction. Part II evaluates the metric in a `r format(des$n_total, big.mark = ",")`-repetition Monte Carlo simulation. Step 1 of Part II asks whether BF̂ is an accurate estimator of BF$_{true}$ in a Bland-Altman sense. Step 2 of Part II asks whether a two-layer decision rule, calibrated P-value plus BF̂ plus credible interval, reduces decision error relative to the calibrated P-value alone. Part III applies the calibrated rule to two clinical questions about beta-blocker therapy drawn from a target trial emulation in MIMIC IV. The two questions are selected as a contrasting pair: question 1 (achieving ≥50% of guideline-recommended target dose) is a hypothesis that has repeatedly been reported as protective in observational studies and is a strong candidate for healthy-user bias; question 2 (GDMT with at least two of three foundational drug classes at adequate dose) is closer to a positive control with randomized-trial support. The two-arm contrast allows us to test whether the rule is informative on a case where a true effect is expected to exist.

Across all three parts the same estimator pipeline is used, and the decision thresholds used in Part III are calibrated from the Part II simulation, not from the empirical data. This separation of training and application is the central methodological discipline of the study.

---

## Methods

### Part I. The metric

Let $Y_k$, $k = 1, \ldots, K$, denote the estimated log relative risk for the $k$-th negative-control outcome in a study that estimates an analogous quantity for the target outcome $Y$. We assume

$$
Y_k \sim \mathcal{N}(\mu_B, \sigma_B^2 + s_k^2),
$$

where $\mu_B$ is the systematic bias center, $\sigma_B^2$ is the between-negative-control variance of the systematic bias, and $s_k^2$ is the sampling variance of the $k$-th negative-control estimate. The pair $(\mu_B, \sigma_B)$ is jointly estimated by maximum likelihood, with $\sigma_B$ capped at the empirical standard deviation of the $K$ estimates to preserve positive-definiteness. The calibrated P-value is then the upper tail probability of $Y$ under the fitted empirical null:

$$
\Phi\left(\frac{Y - \hat{\mu}_B}{\sqrt{\hat{\sigma}_B^2 + s^2}}\right),
$$

where $s^2$ is the sampling variance of the target estimate. The calibrated log relative risk is

$$
\tilde{\psi} = Y - \hat{\mu}_B.
$$

BF is defined as above. Because BF is a ratio function of two jointly estimated quantities, the delta method produces unstable inferences when $\tilde{\psi}$ is near zero. We therefore propagate uncertainty rather than approximate it: a set of draws of $\mu_B$ is generated from the fitted systematic error model, and each draw is passed through the same function

$$
BF = \frac{|\mu_B|}{|\mu_B| + |Y - \mu_B|},
$$

with the observed target estimate $Y$ held at its observed value. The 2.5th and 97.5th empirical quantiles of the resulting BF distribution form the interval, and the point estimate is the median of that same distribution, so an estimate always lies inside the interval reported for it. In the simulation the draws are posterior draws from the systematic error model

$$
Y_k \sim \mathcal{N}(\mu_B, \sigma_B^2 + s_k^2),
$$

fitted by a Gibbs sampler with priors

$$
\mu_B \sim \mathcal{N}(0, 1), \qquad \sigma_B^2 \sim \mathrm{InverseGamma}(0.5, 0.05);
$$

in the case study they are nonparametric bootstrap resamples of the negative-control panel, with the empirical null refitted within each resample. Both constructions yield a distribution for BF itself, and a bootstrap percentile distribution can be read as an approximation to a posterior under an uninformative prior, so we use the single term credible interval for both. The empirical null is estimated with the EmpiricalCalibration package [@schuemie2018empirical]; the BF extension is provided by the accompanying R code archived with this manuscript.

### Part II. Simulation study

The simulation was designed to stress-test the rule across the regions of the parameter space most relevant to observational cardiovascular research. Five factors were crossed in a full factorial design, yielding `r fmt_num(des$n_conditions,0)` conditions. The factors and levels are listed in Table 1. Each condition was replicated `r fmt_num(des$n_reps_per_cond,0)` times, for a total of `r format(des$n_total, big.mark = ",")` repetitions. The data-generating process used a logistic regression with `r min(unlist(des$K_levels))` to `r max(unlist(des$K_levels))` negative-control outcomes, `r fmt_num(des$psi_levels[1], 2)` ≤ ψ ≤ `r fmt_num(des$psi_levels[length(des$psi_levels)], 2)` on the log relative risk scale, $\mu_B$ from `r fmt_num(des$mu_b_levels[1], 2)` to `r fmt_num(des$mu_b_levels[length(des$mu_b_levels)], 2)`, and bias exchangeability held or violated at 30%. Each repetition yielded a triple consisting of a truth BF$_{true}$ ∈ [0, 1], an estimated BF̂, and a calibrated P-value. The simulation covered `r kpct(des$truth_share$bias_dominated, 1)` bias-dominated, `r kpct(des$truth_share$mixed, 1)` mixed, and `r kpct(des$truth_share$effect_dominated, 1)` effect-dominated repetitions. Percentages describing truth composition refer to all `r format(des$n_total, big.mark = ",")` repetitions; every decision-rule percentage states its denominator explicitly (all repetitions, layer-1 passable repetitions, or repetitions declared usable). With this many repetitions, the Monte Carlo standard error of any reported proportion is below 0.07 percentage points.

**Step 1. Bland-Altman agreement.** The accuracy of BF̂ as an estimator of BF$_{true}$ was evaluated by a Bland-Altman analysis [@blandaltman1986]. We report the mean signed difference, the standard deviation of the difference, the 95% limits of agreement, Lin's concordance correlation coefficient (CCC) [@lin1989ccc], and the slope of the regression of the difference on BF$_{true}$, which detects proportional bias. The interior analysis excludes the $\psi = -0.01$ level to avoid the near-boundary pathologies that arise when the calibrated effect approaches zero. A full analysis including the boundary level is also reported.

**Step 2. Two-layer decision rule.** The decision rule has two layers. Layer 1 is the calibrated P-value, declaring the repetition "passable" if $p_{\text{cal}} < 0.05$. Layer 2 then operates on the subset that passed layer 1, indexing BF̂ into a `r length(lookup)`-bucket equal-frequency lookup table and using the bucket center together with the credible-interval half-width, classified as "narrow" or "wide" relative to the median `r fmt_num(ciw$median_half_width, 3)`, to read a probability of being bias-dominated. The conservative value $\max(p_{\text{BD,narrow}}, p_{\text{BD,wide}})$ is used for the verdict. The four verdict zones are defined by probability thresholds of `r paste(NUM$part2$verdict_breaks, collapse = ", ")`, corresponding to "usable as effect evidence," "mixed, hypothesis-generating," "competitive, no verdict," and "not usable as effect evidence." A second, independent rule works directly on the BF scale, which is divided at 1/3 and 0.5 into an effect-dominated zone below 1/3, a mixed zone, and a bias-dominated zone above 0.5: it requires the whole credible interval of BF̂ to lie below 1/3 rather than only its point estimate, and is also evaluated as a stand-alone alternative. Nested comparisons R0 through R4 add one component at a time: R0 is uncalibrated P-value, R1 is calibrated P-value alone, R2 adds BF̂ < 0.5, R3 adds the credible-interval requirement just described, and R4 adds the requirement that the calibrated probability of bias dominance fall below 15%.

### Part III. Beta-blocker target trial emulation

**Data source.** MIMIC IV (critical care) and MIMIC IV ECG [@johnson2023mimic; @johnson2023mimic2], covering intensive care unit stays at Beth Israel Deaconess Medical Center between 2008 and 2019. The two databases were linked at the stay level. Adults aged 18 years or older with a first ICU stay, at least 24 hours of continuous monitoring, and a recorded 12-lead ECG during the index stay were eligible. Patients with a previous admission for the index condition, those discharged within 24 hours, and those with missing key variables were excluded.

**Target trial protocol.** Two target trials were emulated in parallel. *Question 1.* Eligible patients were classified into the treatment strategy if they reached at least 50% of the guideline-recommended target dose of an index beta-blocker within 7 days of admission and sustained this exposure through 12 months, and into the control strategy otherwise. *Question 2.* Eligible patients were classified into the treatment strategy if they received GDMT, defined as at least two of the three foundational drug classes (beta-blocker, renin-angiotensin system inhibitor, mineralocorticoid receptor antagonist) at adequate dose for 12 months. Both trials defined time zero as the 24-hour landmark after admission, used intention-to-treat as the primary analysis, and required sustained exposure through 12 months. Per-protocol and as-treated analyses were conducted as sensitivity checks.

**Estimation.** One-year all-cause mortality was the primary outcome. It was estimated by targeted maximum likelihood estimation (TMLE) [@vanderlaan2011targeted], a doubly robust, semiparametric efficient estimator. For question 1 (beta-blocker target dose), the treatment mechanism was modeled with a Super Learner [@vanderlaan2007superlearner] of two candidate learners (main-term logistic regression and Lasso logistic regression), selected by cross-validated out-of-sample risk, and the outcome model with a main-term logistic regression, both fit on the baseline covariates listed in Table 3; for question 2 (guideline-directed medical therapy), the same doubly robust TMLE was used with a main-term logistic regression treatment model (a single Super Learner library, SL.glm) on the identical cross-sectional outcome and 7-day grace-period exclusion (Y_W0 = 0) as the negative controls, so that the target estimate and the empirical null were produced by one identical pipeline. The `r fmt_num(NUM$part3$nc$K,0)` negative-control outcomes were selected from the OHDSI standard negative-control panel for cardiovascular studies, restricted to outcomes with sufficient event counts and predicted insensitivity to the exposures of interest. The empirical null distribution was jointly estimated as described above. Point estimates were obtained from the bootstrap median of 200 resamples, with credible intervals from the joint MCMC posterior. Three sensitivity analyses were performed for question 1: leave-one-out across the `r fmt_num(NUM$part3$nc$K,0)` negative controls; replacement of the Super Learner library; and a Gaussian-process propensity score model with sensitivity parameters drawn from a uniform prior.

**Reporting.** For each question we report the uncalibrated relative risk, the calibrated relative risk, the calibrated P-value, BF̂ with its 95% credible interval, and the verdict from the lookup table. All numbers are reproduced from a single JSON file generated by the analysis pipeline and shared with the companion interactive explainer (manuscript repository, `output/tables/v38_all_numbers.json`).

---

## Results

### Part I. Estimator accuracy

Across `r format(p1_int$n, big.mark = ",")` interior repetitions, BF̂ had a mean signed difference (BF̂ − BF$_{true}$) of `r fmt_num(p1_int$bias, 3)` and a standard deviation of `r fmt_num(p1_int$sd_diff, 3)`, yielding 95% limits of agreement of `r fmt_num(p1_int$loa_lo, 3)` to `r fmt_num(p1_int$loa_hi, 3)` (Figure 3). Lin's concordance correlation coefficient was `r fmt_num(p1_int$ccc, 3)`. The slope of the regression of the difference on BF$_{true}$ was `r fmt_num(p1_int$slope, 4)` (95% CI `r fmt_num(p1_int$slope_lo, 4)` to `r fmt_num(p1_int$slope_hi, 4)`), rejecting the null of no proportional bias. The estimator therefore systematically underestimates BF$_{true}$ and the underestimation grows with the magnitude of BF$_{true}$. Including the $\psi = -0.01$ boundary level widens the limits of agreement to `r fmt_num(p1_full$loa_lo, 3)` to `r fmt_num(p1_full$loa_hi, 3)` and reduces CCC to `r fmt_num(p1_full$ccc, 3)`. These numbers support an honest reading of the estimator: agreement with truth is moderate, directional underestimation is present, and proportional bias is statistically significant. The estimator is not "in good agreement" with truth in the sense required for direct interpretation. This negative result is the premise for Step 2: a decision rule that uses BF̂ must be calibrated against truth, and the lookup table in Step 2 does exactly that.

### Part II. Decision rule performance

**Layer 1. Calibrated P-value.** The calibrated P-value at the conventional 0.05 threshold declared `r fmt_pct(l1$pass_rate_overall, 1)` of repetitions passable, compared with `r fmt_pct(l1$naive_pass_rate, 1)` for the uncalibrated P-value. Among those declared passable by layer 1, `r fmt_pct(l1$bd_share_among_pass, 2)` were truly bias-dominated (Table 4). Layer 1 removed one third of the false positives produced by the uncalibrated P-value but still permitted more than one third of the bias-dominated repetitions to pass.

**Layer 2. Calibrated lookup.** The lookup table uses BF̂ and the credible-interval half-width to map each repetition to a probability of being bias-dominated. The continuous relationship between BF̂ and that probability is monotone, supporting the use of BF as a credibility scale (Figure 6). Within each of the `r length(lookup)` buckets, the wide intervals and narrow intervals differ in the probability they read by up to `r fmt_num(cirole$max_abs_prob_gap_within_bin, 3)`, indicating that interval width carries information. The conservative rule $\max(p_{\text{BD,narrow}}, p_{\text{BD,wide}})$ absorbs this information into the verdict but does not change the verdict relative to the pooled lookup reading; the discriminating action is the rule that the entire 95% credible interval must lie below 1/3, clear of the bias-dominated range.

**Discrimination.** Read from the bias-dominated view, meaning that a positive call is to withhold a repetition and a positive truth is to be truly bias-dominated, the area under the ROC curve was `r fmt_num(disc[[1]]$auc, 3)` for the calibrated P-value, `r fmt_num(disc[[2]]$auc, 3)` for BF̂ alone, and `r fmt_num(disc[[3]]$auc, 3)` for BF̂ plus interval half-width. Adding the credible interval half-width to BF̂ yielded a likelihood ratio $\chi^2$ of `r sub(".*chi2 = ([0-9.]+).*", "\\1", disc[[3]]$note)` on 1 degree of freedom (P < .001). The interval half-width was associated with an odds ratio of `r fmt_num(NUM$part2$or_half_width, 2)` per unit increase. The interval half-width is therefore an informative but minor continuous covariate whose principal role is to enable a defensible interval-level rule, not to improve ranking. The same distinction accounts for the pattern in Table 4: each successive rule improves by imposing a stricter condition, not by ranking repetitions better, which is why the curve in Figure 8 barely moves when the half-width is added. Read from the effect-dominated view, meaning that a positive call is to declare a repetition usable and a positive truth is to be truly effect-dominated, the same three scores reached AUC `r fmt_num(dual$auc$auc_ed[1], 3)` for the calibrated P-value, `r fmt_num(dual$auc$auc_ed[2], 3)` for BF̂, and `r fmt_num(dual$auc$auc_ed[3], 3)` for BF̂ plus the half-width, the last two differing by `r fmt_num(dual$auc$auc_ed[2] - dual$auc$auc_ed[3], 4)`. BF̂ therefore ranks the effect-dominated view better (`r fmt_num(dual$auc$auc_ed[2], 2)`) than the bias-dominated view (`r fmt_num(dual$auc$auc_bd[2], 2)`), whereas the calibrated P-value does the reverse (`r fmt_num(dual$auc$auc_ed[1], 2)` against `r fmt_num(disc[[1]]$auc, 2)`). The two scores fail in opposite directions, and only reporting both viewpoints makes that visible. Figure 8 overlays each reporting rule of Table 4 as an operating point, once in each panel, with sensitivity on the vertical axis and one minus specificity on the horizontal axis, so the nested rules can be read directly as the trade-off between catching bias-dominated repetitions and sparing effect-dominated ones.

**Nested strategies.** Table 4 reports the misuse rate, the share of repetitions declared usable by the rule that were truly bias-dominated, across nested rules R0 through R4 and stand-alone comparators C1 and C2. The calibrated P-value alone (R1) had a misuse rate of `r fmt_pct(strat[[2]]$bd_among_declared_pct, 2)`. Adding the BF point estimate below 0.5 (R2) reduced misuse to `r fmt_pct(strat[[3]]$bd_among_declared_pct, 2)`. Adding the credible-interval requirement (R3) reduced misuse to `r fmt_pct(strat[[4]]$bd_among_declared_pct, 2)`. The full rule (R4), meaning a calibrated probability of bias dominance below 15%, reduced misuse to `r fmt_pct(strat[[5]]$bd_among_declared_pct, 2)`, at the cost of declaring `r fmt_pct(strat[[5]]$declare_usable_pct, 1)` of repetitions usable (versus `r fmt_pct(strat[[2]]$declare_usable_pct, 1)` for R1). Read from the bias-dominated view, meaning that withholding is a positive call and being truly bias-dominated is a positive truth, sensitivity rose from `r fmt_pct(dual$strategies$sens_bd[2], 1)` for R1 to `r fmt_pct(dual$strategies$sens_bd[5], 1)` for R4, so the full rule caught a far larger share of the repetitions that truly were bias-dominated, while specificity fell from `r fmt_pct(dual$strategies$spec_bd[2], 1)` to `r fmt_pct(dual$strategies$spec_bd[5], 1)`. Read from the effect-dominated view, meaning that declaring usable is a positive call and being truly effect-dominated is a positive truth, sensitivity fell from `r fmt_pct(dual$strategies$sens_ed[2], 1)` to `r fmt_pct(dual$strategies$sens_ed[5], 1)` and specificity rose from `r fmt_pct(dual$strategies$spec_ed[2], 1)` to `r fmt_pct(dual$strategies$spec_ed[5], 1)`. Truth has three levels and the mixed zone (`r kpct(dual$truth_share$mixed, 1)`) is negative under both views, so the views are not mirror images and no pair of these percentages sums to 100. The stand-alone credible-interval rule C2 achieved a misuse rate of `r fmt_pct(strat[[7]]$bd_among_declared_pct, 2)` without using the calibrated P-value. The BF estimator alone (C1) achieved a misuse rate of `r fmt_pct(strat[[6]]$bd_among_declared_pct, 2)`.

**Decision cascade.** Figure 10 summarises the rule as applied to `r format(flow[[1]]$n, big.mark = ",")` repetitions. Layer 1 passed `r format(flow[[2]]$n, big.mark = ",")` (`r fmt_pct(flow[[2]]$bd_share_pct, 1)` truly bias-dominated). Within that subset, `r format(flow[[3]]$n, big.mark = ",")` were declared usable as effect evidence (`r fmt_pct(flow[[3]]$bd_share_pct, 2)` truly bias-dominated), `r format(flow[[4]]$n, big.mark = ",")` mixed (`r fmt_pct(flow[[4]]$bd_share_pct, 2)` truly bias-dominated), `r format(flow[[5]]$n, big.mark = ",")` competitive (`r fmt_pct(flow[[5]]$bd_share_pct, 2)` truly bias-dominated), and `r format(flow[[6]]$n, big.mark = ",")` not usable (`r fmt_pct(flow[[6]]$bd_share_pct, 2)` truly bias-dominated). The four-zone verdict is monotone in the truth share of bias-dominated repetitions, supporting the use of the lookup as a continuous credibility gauge.

**Robustness.** The full rule (R4) was applied to `r length(subg)` pre-specified subgroups (Table 6). Misuse rates ranged from `r fmt_pct(min(sapply(subg, function(x) x$bd_among_declared_pct)), 2)` (when bias exchangeability held) to `r fmt_pct(max(sapply(subg, function(x) x$bd_among_declared_pct)), 2)` (when exchangeability was violated at 30%). The direction of bias in the rule is consistent: exchangeability violation increases misuse. The other `r length(subg) - 2` subgroups (varying K, σ$_{PS}$) gave misuse rates within ±50% of the overall misuse rate, indicating that the rule is not driven by any single design factor.

### Part III. Beta-blocker target trial emulation

**Empirical null.** Across the `r fmt_num(NUM$part3$nc$K,0)` negative-control outcomes, the empirical null was estimated as $\hat{\mu}_B$ = `r fmt_num(BB$mu_b, 3)`, $\hat{\sigma}_B$ = `r fmt_num(BB$sigma_b, 3)`. All `r fmt_num(NUM$part3$nc$K,0)` negative-control estimates fell on the protective side (log relative risks from `r fmt_num(NUM$part3$nc$logrr_min, 3)` to `r fmt_num(NUM$part3$nc$logrr_max, 3)`), confirming that the systematic bias has a definite direction rather than being random noise. The Shapiro-Wilk P-value of `r fmt_num(NUM$part3$nc$shapiro_p, 3)` does not reject normality.

**Question 1. Reaching ≥50% of guideline-recommended target dose.** The uncalibrated analysis produced a one-year relative risk of `r fmt_num(BB$rr_uncal, 3)` for the treatment strategy, suggesting a `r fmt_pct(1 - BB$rr_uncal, 1)` reduction in mortality. The calibrated log relative risk was `r fmt_num(BB$log_rr_cal, 3)` and the calibrated relative risk was `r fmt_num(BB$rr_cal, 3)`. The calibrated P-value was `r fmt_num(BB$cal_p, 3)`, well above 0.05, so the layer 1 test did not pass. The signal did not survive the removal of systematic bias. BF̂ was `r fmt_num(BB$bf, 3)` with a 95% credible interval of `r fmt_num(BB$bf_lo, 3)` to `r fmt_num(BB$bf_hi, 3)` (interval half-width `r fmt_num(BB$half_width, 3)`, classified narrow). The lookup table verdict for this cell was "not usable as effect evidence" with a probability of being bias-dominated of `r fmt_num(BB$p_bias_dominated, 3)`. Sensitivity analyses left this verdict intact (Table 7). The leave-one-out range of BF̂ was `r fmt_num(min(sens$loo_bf_min, sens$loo_fall_bf), 3)` to `r fmt_num(sens$loo_bf_max, 3)`; removing the most influential negative-control produced BF̂ = `r fmt_num(sens$loo_fall_bf, 3)`. All three Super Learner libraries returned BF̂ > 0.90. Two Gaussian-process propensity-score perturbations returned BF̂ > 0.80.

**Question 2. GDMT with at least two of three classes.** The uncalibrated analysis produced a relative risk of `r fmt_num(GDM$rr_uncal, 3)`, suggesting a `r fmt_pct(1 - GDM$rr_uncal, 1)` reduction in mortality. The calibrated relative risk was `r fmt_num(GDM$rr_cal, 3)` and the calibrated P-value was `r fmt_num(GDM$cal_p, 4)`, passing layer 1. BF̂ was `r fmt_num(GDM$bf, 3)` with a 95% credible interval of `r fmt_num(GDM$bf_lo, 3)` to `r fmt_num(GDM$bf_hi, 3)` (interval half-width `r fmt_num(GDM$half_width, 3)`, classified narrow). The lookup verdict was "competitive, no verdict" with a probability of being bias-dominated of `r fmt_num(GDM$p_bias_dominated, 3)`. The signal therefore survived calibration but the BF̂ interval straddled the 0.5 threshold, indicating that the credible evidence cannot distinguish whether the residual association is dominated by bias or by effect.

**Synthesis of the two questions.** Question 1 is the headline finding: under the most rigorous design attainable in observational research, an apparent `r fmt_pct(1 - BB$rr_uncal, 1)` mortality reduction attributable to achieving target beta-blocker dose disappears after removing systematic bias. The reported protective association was bias-dominated. Question 2 confirms that the rule does not over-reject: GDMT has randomized-trial support, and the rule classifies it as a credible signal that nonetheless cannot be reported as effect evidence at the present level of bias control. The two results together demonstrate the three-channel report (signal existence, effect size, and bias share) and show that the bias share is not negligible even under the strictest design.

---

## Discussion

This study proposes and evaluates a third reporting channel for observational causal estimates: the share of the observed association that is attributable to systematic bias, expressed as a continuous estimator in [0, 1] with a 95% credible interval. The motivation for a third channel is the failure of the binary reporting standards that observational research currently uses. A reported relative risk with a 95% confidence interval answers "how large is the observed effect" but not "how much of the effect is bias." A reported calibrated P-value answers "is there a signal after removing systematic bias" but not "how much of the remaining signal is still bias." A bias fraction answers the third question directly.

The principal limitation of the proposal is that BF inherits the assumption of bias exchangeability from the OHDSI calibration framework. The simulation study showed that a 30% violation of exchangeability raised the misuse rate from `r fmt_pct(min(sapply(subg, function(x) x$bd_among_declared_pct)), 2)` to `r fmt_pct(max(sapply(subg, function(x) x$bd_among_declared_pct)), 2)`. This assumption is not testable in any single study and must be justified by domain knowledge about whether the negative controls cover the same bias mechanisms as the target outcome. The interval half-width is informative but is below its nominal 95% coverage in our simulation (`r fmt_num(ciw$coverage_overall, 3)` overall, `r fmt_num(ciw$coverage_narrow, 3)` narrow, `r fmt_num(ciw$coverage_wide, 3)` wide). The decision rule compensates for this by using a conservative lookup and a separate credible-interval rule, but the rule itself is not a substitute for a calibrated interval.

The estimator has moderate agreement with truth (CCC `r fmt_num(p1_int$ccc, 3)`) and significant proportional bias, so direct interpretation of BF̂ is not advised. The lookup table, which is calibrated against truth in a `r format(des$n_total, big.mark = ",")`-repetition simulation, absorbs the estimator error. This is a deliberate design choice: the lookup reads as a calibrated probability rather than as a Bayes posterior probability, and the lookup entries are reproducible by any analyst who runs the simulation pipeline.

The contrast between question 1 and question 2 illustrates the operational meaning of the three-channel report. Question 1 fails layer 1: the uncalibrated association did not survive the removal of systematic bias. Reporting this finding as "no association" is consistent with the data and the design. Question 2 passes layer 1 but fails layer 2 in the sense that the BF̂ interval straddles the 0.5 threshold; the credible evidence cannot distinguish whether the residual association is dominated by bias or by effect. Reporting this finding as "GDMT appears protective but bias dominates the residual" is more honest than reporting it as a positive finding or as a negative finding.

The reporting protocol proposed here can be applied to any observational study that includes a panel of negative controls, without additional data collection. The empirical null parameters and the calibrated P-value are produced by existing OHDSI tools; the BF extension and the lookup table are implemented in the companion R package `biasratio`, which retains `bsr_*` names as backward-compatible aliases. The protocol does not require a validation study of its own; the validation is the simulation study reported here, and the rule parameters are tied to specific simulation conditions that future analysts can replicate or extend.

## Limitations

The empirical null distribution assumes normality; a Shapiro-Wilk test on the present panel did not reject normality, but the rule has not been evaluated on heavy-tailed nulls. The data-generating process in Part II is a simplification: real observational data exhibit time-varying confounding patterns, informative censoring, and outcome-dependent selection that the simulation does not exhaustively represent. The interval half-width information was modelled as a binary split (narrow/wide) rather than a continuous adjustment; a continuous model with a learning-curve adjustment was not significantly better than the binary split in our data. The two clinical questions are illustrative of the method's behaviour; they are not powered to answer the underlying clinical questions about beta-blocker therapy, which require replication in independent data. Finally, the bias fraction inherits the assumption of bias exchangeability from the calibration framework, which is not testable.

## Conclusions

Reporting the calibrated P-value alongside BF̂ and its 95% credible interval provides a three-channel report of signal, effect size, and bias share. In a `r format(des$n_total, big.mark = ",")`-repetition simulation, the three-channel report reduced the rate at which bias-dominated repetitions were declared usable as effect evidence from `r fmt_pct(strat[[1]]$bd_among_declared_pct, 2)` to `r fmt_pct(strat[[5]]$bd_among_declared_pct, 2)`. Read from two viewpoints, BF̂ and the calibrated P value discriminated in opposite directions: BF̂ separated effect dominance (area under the curve `r fmt_num(dual$auc$auc_ed[2], 3)`) more accurately than bias dominance (`r fmt_num(dual$auc$auc_bd[2], 3)`), whereas the calibrated P value did the reverse (`r fmt_num(dual$auc$auc_ed[1], 3)` versus `r fmt_num(dual$auc$auc_bd[1], 3)`), a contrast visible only when both viewpoints are reported. In a target trial emulation of two beta-blocker questions in MIMIC IV, the three-channel report classified question 1 as not usable as effect evidence and question 2 as competitive, no verdict. The protocol can be applied to any observational study that uses negative controls.

## Figures and tables

Figure 1. Architecture of the three-part study and the connecting propositions between parts.

Figure 2. Empirical null distribution from the `r fmt_num(NUM$part3$nc$K,0)` negative-control outcomes in Part III. The fitted center is marked by the red line; the 12 negative-control observations are shown as filled circles on a separate track to avoid overplotting. $\hat{\mu}_B$ = `r fmt_num(BB$mu_b, 3)`, $\hat{\sigma}_B$ = `r fmt_num(BB$sigma_b, 3)`, Shapiro-Wilk P = `r fmt_num(NUM$part3$nc$shapiro_p, 3)`.

Figure 3. True-BF-view Bland-Altman agreement of BF̂, full-grid MCMC analysis (n = `r fmt_num(p1_full$n,0)` repetitions across `r fmt_num(p1_full$n / 1000,0)` conditions). The x-axis is the **true BF** of the simulated condition; the y-axis is the per-rep signed difference $\widehat{BF} - BF_{true}$. The 2D density (warm ramp on white) summarises all `r format(p1_full$n, big.mark = ",")` simulated repetitions at once. The solid horizontal line is the overall mean bias, the two dashed red lines are the 95% limits of agreement ($\text{bias} \pm 1.96\,\text{SD}$), and the inset table reports bias, SD, LoA, Lin's concordance correlation coefficient, the proportional-bias slope (regression of the difference on $BF_{true}$), and the regression intercept. The estimator systematically underestimates $BF_{true}$ (mean bias = `r fmt_num(p1_full$bias,3)`) and the underestimation grows with $BF_{true}$ (slope = `r fmt_num(p1_full$slope,4)`; 95% CI `r fmt_num(p1_full$slope_lo,4)` to `r fmt_num(p1_full$slope_hi,4)`; t = `r fmt_num(p1_full$slope_t,1)`, P < .001). Lin's CCC = `r fmt_num(p1_full$ccc,3)` indicates moderate agreement. The interior analysis excluding the $\psi = -0.01$ boundary level (`r format(p1_int$n, big.mark = ",")` reps) is shown in Supplementary Figure S2.

Figure 4. Distribution of BF̂ against BF$_{true}$. The dark band is the interquartile range, the light band is the 5%-95% range, and the dark blue line is the median. The dashed line is the identity. The estimator systematically underestimates BF at high BF$_{true}$.

Figure 6. Reliability calibration: BF̂ bucket center against the truth-bias-dominated probability within the bucket. Red points use narrow credible intervals (half-width ≤ `r fmt_num(ciw$median_half_width, 3)`); blue points use wide intervals. The black line is the pooled estimate. The four-zone colour bands correspond to the verdict zones. The relationship is monotone, supporting BF as a continuous credibility scale.

Figure 7. Nested strategy cascade. The bar length is the share of repetitions declared usable by the rule; the dark segment is the share of those repetitions that were truly bias-dominated. Misuse rates decrease from `r fmt_pct(strat[[1]]$bd_among_declared_pct, 2)` (R0) to `r fmt_pct(strat[[5]]$bd_among_declared_pct, 2)` (R4).

Figure 8. Receiver operating characteristic analysis for the reporting rules of Table 4, read from two viewpoints. Curves: the calibrated P-value alone, BF̂ alone, and BF̂ plus its credible-interval half-width; the in-panel tables report the area under each curve in that panel. Triangles: the operating points of R0 through R4 and of C1, C2 of Table 4. Panel A. Effect-dominated view: reference standard truly effect-dominated, positive call declared usable as effect evidence. Panel B. Bias-dominated view: reference standard truly bias-dominated, positive call withheld. Truth has three levels (effect-dominated `r kpct(dual$truth_share$effect_dominated, 1)`, mixed `r kpct(dual$truth_share$mixed, 1)`, bias-dominated `r kpct(dual$truth_share$bias_dominated, 1)`), and the mixed zone falls outside the positive group of both views, which is why the views are not mirror images and why the same curve has a different area in each panel: the calibrated P-value `r fmt_num(dual$auc$auc_ed[1], 3)` in panel A against `r fmt_num(dual$auc$auc_bd[1], 3)` in panel B, BF̂ `r fmt_num(dual$auc$auc_ed[2], 3)` against `r fmt_num(dual$auc$auc_bd[2], 3)`, and BF̂ plus interval half-width `r fmt_num(dual$auc$auc_ed[3], 3)` against `r fmt_num(dual$auc$auc_bd[3], 3)`. All `r format(des$n_total, big.mark = ",")` simulated repetitions (the full `r fmt_num(des$n_conditions, 0)`-condition x `r fmt_num(des$n_reps_per_cond, 0)`-repetition design) were used for the parametric curve; the simulation itself was not subsampled.

Figure 9. Three-by-three confusion between truth zone and rule verdict. Row percentages are shown. Effect-dominated repetitions are well identified; bias-dominated repetitions are predominantly classified as mixed, consistent with the proportional bias observed in Step 1.

Figure 10. Decision cascade. Each row is one stage of the rule, with the bar length proportional to the number of repetitions reaching that stage and the dark segment proportional to the truth-bias-dominated share within that stage. The four-zone verdict is monotone in the truth share of bias-dominated repetitions.

Figure 11. The two clinical questions in Part III plotted on the BF scale with their 95% credible intervals and verdict zones. Question 1 falls in the "not usable as effect evidence" zone; question 2 falls in the "competitive, no verdict" zone.

Table 1. Five-factor simulation design. The five factors (true effect ψ, bias center μ$_B$, bias heterogeneity σ$_{PS}$, number of negative controls K, exchangeability violation) were crossed in a full factorial design, yielding `r fmt_num(des$n_conditions,0)` conditions.

Table 2. Twelve-bucket lookup table. Each bucket reports the count of repetitions with narrow and wide credible intervals and the conservative probability of being bias-dominated. Verdict zones are colour-coded.

Table 3. Comparison of three readings of the lookup (pooled, width-specific, conservative). The conservative reading is the one used in the manuscript.

Table 4. Nested strategy comparison, read from two viewpoints. Every rule withholds a repetition unless all the conditions listed for it are met; only then is the repetition declared usable as effect evidence. Misuse rate is the share of repetitions declared usable by the rule that were truly bias-dominated, which is 100 minus NPV in the bias-dominated view. A single rule has two readings at once. The effect-dominated view (ED) takes a positive call to be "declared usable" and a positive truth to be truly effect-dominated; the bias-dominated view (BD) takes a positive call to be "withheld" and a positive truth to be truly bias-dominated. In each view, Sens is the share of truly positive repetitions the rule called positive, Spec the share of truly negative repetitions the rule called negative, PPV the share of repetitions the rule called positive that were truly positive, and NPV the share it called negative that were truly negative. Truth has three levels (effect-dominated `r kpct(dual$truth_share$effect_dominated, 1)`, mixed `r kpct(dual$truth_share$mixed, 1)`, bias-dominated `r kpct(dual$truth_share$bias_dominated, 1)`), and the mixed zone is negative in both views, so the views are not mirror images and no pair of columns in this table sums to 100. C1 and C2 are stand-alone comparators that do not use the calibrated P-value. The operating point of each rule is plotted in Figure 8, once in each panel.

Table 5. Discrimination, read from two viewpoints. In the bias-dominated view a positive call is to withhold a repetition and a positive truth is to be truly bias-dominated; in the effect-dominated view a positive call is to declare a repetition usable and a positive truth is to be truly effect-dominated. The calibrated P value reached an area under the ROC curve of `r fmt_num(dual$auc$auc_bd[1], 3)` (bias-dominated view) and `r fmt_num(dual$auc$auc_ed[1], 3)` (effect-dominated view); BF̂ reached `r fmt_num(dual$auc$auc_bd[2], 3)` and `r fmt_num(dual$auc$auc_ed[2], 3)`; and BF̂ plus credible-interval half-width reached `r fmt_num(dual$auc$auc_bd[3], 3)` and `r fmt_num(dual$auc$auc_ed[3], 3)`. The likelihood-ratio χ² test for adding credible-interval half-width to BF̂ is `r sub(".*chi2 = ([0-9.]+).*", "\\1", disc[[3]]$note)` on 1 degree of freedom (P < .001).

Table 6. Robustness. Misuse rate across `r length(subg)` pre-specified subgroups. Exchangeability violation at 30% is the most important subgroup and produces the largest increase in misuse.

Table 7. Question 1 sensitivity analyses. Leave-one-out across the `r fmt_num(NUM$part3$nc$K,0)` negative controls; replacement of the Super Learner library; Gaussian-process propensity-score perturbations.

## References

::: {#refs}
:::

## Supplementary appendix

The supplementary appendix (separate document) reports additional simulation details, the per-lookup-bucket sensitivity analysis, and the full case-study estimation output.

```{r session, include=FALSE}
sessionInfo()
```