---
title: "Supplementary Appendix: Quantifying Systematic Bias in Observational Causal Estimates Using Negative Controls: The Bias Fraction"
author:
  - name: Zeng Zhechun
    degrees: MD, MPH
    affiliation: "1,2"
    email: zengzhechun@icloud.com
    corresponding: true
  - name: Wang Jinwen
    degrees: MD, PhD
    affiliation: "1,2"
  - name: Zuo Huijuan
    degrees: MD, MPH
    affiliation: "1,2"
  - name: Shu Lixia
    degrees: PhD
    affiliation: "1,2"
date: "2026-08-27"
bibliography: references.bib
csl: american-medical-association.csl
format:
  docx:
    toc: false
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
---

^1^ Beijing Anzhen Hospital, Capital Medical University, Beijing, China

^2^ Beijing Institute of Heart, Lung and Blood Vessel Diseases, Beijing, China

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(jsonlite)
  library(knitr)
})
source("../R/00_config.R")
source("../R/01_bsr_core.R")
library(EmpiricalCalibration)

NUM <- jsonlite::fromJSON(file.path(TAB_DIR, "v38_all_numbers.json"),
                          simplifyVector = FALSE)

# Dual-view operating characteristics: the same rules R0-R4, C1-C2 scored twice,
# once taking "declared usable" as the positive call (effect-dominated view, ED)
# and once taking "withheld" as the positive call (bias-dominated view, BD), each
# with sensitivity, specificity, PPV and NPV. Written by R/23_dual_view_metrics.R.
dual <- jsonlite::fromJSON(file.path(TAB_DIR, "v38_dual_view.json"),
                           simplifyVector = TRUE)

des       <- NUM$design
p1_int    <- NUM$part1$interior
p1_full   <- NUM$part1$full
l1        <- NUM$part2$layer1
ciw       <- NUM$part2$ci_width
disc      <- NUM$part2$discrimination
strat     <- NUM$part2$strategies
lookup    <- NUM$part2$lookup
flow      <- NUM$part2$flow
subg      <- NUM$part2$subgroups
cirole    <- NUM$part2$ci_width_role
confusion3x3 <- NUM$part2$confusion3x3
cases     <- NUM$part3$cases
sens      <- NUM$part3$sensitivity
nc        <- NUM$part3$nc

fmt_pct <- function(x, d = 1) sprintf(paste0("%.", d, "f"), as.numeric(x))
fmt_num <- function(x, d = 2) sprintf(paste0("%.", d, "f"), as.numeric(x))
# kpct is for fields stored as proportions (eg dual$truth_share$mixed = 0.2875)
# that must be shown as 28.7%.
kpct <- function(x, d = 1) paste0(fmt_pct(100 * as.numeric(x), d), "%")
# helper to make a kable with rounded percentage / numeric columns
kpct <- function(x, d = 1) paste0(fmt_pct(100 * as.numeric(x), d), "%")
```

This supplementary appendix provides the methodological detail, full lookup table, and complete case-study estimation output that support the main manuscript. All numbers are reproduced from the single JSON file `output/tables/v38_all_numbers.json` generated by the analysis pipeline, which is shared with the companion interactive explainer.

---

## S1. Extended methods

### S1.1 Full simulation design

The simulation crossed five factors in a full factorial design, yielding `r fmt_num(des$n_conditions, 0)` conditions. Each condition was replicated `r fmt_num(des$n_reps_per_cond, 0)` times for a total of `r format(des$n_total, big.mark = ",")` repetitions. The factors and levels were:

- True effect on the log relative risk scale, ψ, with `r length(des$psi_levels)` levels from `r fmt_num(des$psi_levels[1], 2)` to `r fmt_num(des$psi_levels[length(des$psi_levels)], 2)`.
- Systematic bias center, μ_B, with `r length(des$mu_b_levels)` levels from `r fmt_num(des$mu_b_levels[1], 2)` to `r fmt_num(des$mu_b_levels[length(des$mu_b_levels)], 2)`.
- Between-negative-control variance of the bias, σ_ps, with `r length(des$sigma_ps_levels)` levels (`r fmt_num(des$sigma_ps_levels[1], 2)`, `r fmt_num(des$sigma_ps_levels[2], 2)`).
- Number of negative controls, K, with `r length(des$K_levels)` levels (`r paste(des$K_levels, collapse = ", ")`).
- Exchangeability violation, with `r length(des$exV_levels)` levels (`r des$exV_levels[1]`, `r des$exV_levels[2]`, where 0.3 denotes a 30% violation rate).

The data-generating process was a logistic regression in which each negative-control estimate was drawn as Y_k ~ N(μ_B, σ_B^2 + s_k^2), with the target estimate generated under the analogous model. By construction, each repetition carried a known truth BF_{true} in [0, 1], an estimated BF̂, and a calibrated P-value. The design allocated `r format(des$n_interior, big.mark = ",")` interior repetitions (excluding the ψ = -0.01 boundary level) and `r format(des$n_boundary, big.mark = ",")` boundary repetitions. The truth composition was `r kpct(des$truth_share$bias_dominated, 1)` bias-dominated, `r kpct(des$truth_share$mixed, 1)` mixed, and `r kpct(des$truth_share$effect_dominated, 1)` effect-dominated.

### S1.2 Bias Fraction definition and interval construction

The Bias Fraction is BF = |μ̂_B| / (|μ̂_B| + |ψ̃|), where μ̂_B is the estimated center of the empirical null and ψ̃ is the calibrated log relative risk. Because BF is a ratio of two jointly estimated quantities, the delta method is unstable when ψ̃ is near zero. The credible interval was therefore constructed by propagating uncertainty rather than approximating it: a set of draws of μ_B was generated from the fitted systematic error model, and each draw was passed through the same function BF = |μ_B| / (|μ_B| + |Y - μ_B|), with the observed target estimate Y held at its observed value. The 2.5% and 97.5% empirical quantiles of the resulting BF distribution formed the interval, and the point estimate was the median of that same distribution, so an estimate always lies inside the interval reported for it. In the simulation the draws were posterior draws from the systematic error model Y_k ~ N(μ_B, σ_B^2 + s_k^2), fitted by a Gibbs sampler with priors μ_B ~ N(0, 1) and σ_B^2 ~ InverseGamma(0.5, 0.05); in the case study they were nonparametric bootstrap resamples of the negative-control panel, with the empirical null refitted within each resample. Both constructions yield a distribution for BF itself, and a bootstrap percentile distribution can be read as an approximation to a posterior under an uninformative prior, so a single term, credible interval, is used for both.

### S1.3 Negative-control selection for the case study

For the beta-blocker target trial emulation, `r fmt_num(nc$K, 0)` negative-control outcomes were drawn from the OHDSI standard negative-control panel for cardiovascular studies, restricted to outcomes with sufficient event counts and predicted insensitivity to the exposures of interest. All `r fmt_num(nc$K, 0)` selected controls were directionally protective (all_protective = `r nc$all_protective`), with calibrated log relative risks ranging from `r fmt_num(nc$logrr_min, 3)` to `r fmt_num(nc$logrr_max, 3)` and a Shapiro-Wilk test of the empirical null residuals yielding p = `r fmt_num(nc$shapiro_p, 3)`, consistent with approximate normality of the null.

---

## S2. Part I supplementary: Bland-Altman agreement

The interior analysis (primary) excludes the ψ = -0.01 boundary level to avoid near-boundary pathologies. The full analysis includes it. Both are reported below.

```{r ba-table}
ba <- data.frame(
  Analysis = c("Interior (primary)", "Full (incl. boundary)"),
  N = c(p1_int$n, p1_full$n),
  `Mean diff (BFhat-BFtrue)` = c(p1_int$bias, p1_full$bias),
  `SD diff` = c(p1_int$sd_diff, p1_full$sd_diff),
  `LoA low` = c(p1_int$loa_lo, p1_full$loa_lo),
  `LoA high` = c(p1_int$loa_hi, p1_full$loa_hi),
  `CCC` = c(p1_int$ccc, p1_full$ccc),
  `Prop. bias slope` = c(p1_int$slope, p1_full$slope),
  MAE = c(p1_int$mae, p1_full$mae),
  RMSE = c(p1_int$rmse, p1_full$rmse)
)
kable(ba, digits = c(0, 0, 3, 3, 3, 3, 3, 4, 3, 3),
      caption = "Table S1. Bland-Altman agreement of BFhat as an estimator of BFtrue.")
```

The worst single-condition mean signed difference among the `r fmt_num(NUM$part1$n_conditions_interior, 0)` interior conditions was `r fmt_num(NUM$part1$interior_worst_cond_bias, 3)`. The proportional-bias slope regression rejected the null of no proportional bias in both analyses (interior slope `r fmt_num(p1_int$slope, 4)`, 95% CI `r fmt_num(p1_int$slope_lo, 4)` to `r fmt_num(p1_int$slope_hi, 4)`; full slope `r fmt_num(p1_full$slope, 4)`, 95% CI `r fmt_num(p1_full$slope_lo, 4)` to `r fmt_num(p1_full$slope_hi, 4)`). The estimator systematically underestimates BF_{true}, and the underestimation grows with the magnitude of BF_{true}.

---

## S3. Part II supplementary: decision-rule detail

### S3.1 Full 12-bucket lookup table

Each bucket reports the count of repetitions with narrow and wide credible intervals, the probability of being bias-dominated under the narrow reading, the wide reading, and the conservative max(narrow, wide) reading, together with the verdict zone.

```{r lookup-table}
lk_rows <- function(L) {
  data.frame(
    Bin = sapply(L, `[[`, "bin"),
    `BF centre` = sapply(L, `[[`, "bf_centre"),
    `n narrow` = sapply(L, `[[`, "n_narrow"),
    `n wide` = sapply(L, `[[`, "n_wide"),
    `pBD narrow` = sapply(L, function(x) kpct(x$p_bd_narrow, 2)),
    `pBD wide` = sapply(L, function(x) kpct(x$p_bd_wide, 2)),
    `pBD cons.` = sapply(L, function(x) kpct(x$p_bd_conservative, 2)),
    Verdict = sapply(L, `[[`, "verdict")
  )
}
kable(lk_rows(lookup), digits = c(0, 3, 0, 0, 0, 0, 0, 0),
      caption = "Table S2. Twelve-bucket equal-frequency lookup table mapping BFhat x interval width to probability of being bias-dominated.")
```

### S3.2 Three readings of the lookup

The lookup supports a pooled reading (CI width ignored), a width-specific reading, and the conservative max(narrow, wide) reading used in the manuscript.

```{r reading-table}
rr <- cirole$rules
rd <- data.frame(
  Reading = sapply(rr, `[[`, "reading"),
  `Misuse rate (%)` = sapply(rr, function(x) fmt_pct(x$bd_among_declared_pct, 2)),
  `Declare usable (%)` = sapply(rr, function(x) fmt_pct(x$declare_usable_pct, 1)),
  # spec_bd_pct 是 R/18 的历史命名，它算的是 P(扣下 | 真偏倚主导)，即 BD 视角的
  # 灵敏度而非特异度；表头按正确语义写，变量名保持不变（改变量名要动 R/18）。
  `Sens ED (%)` = sapply(rr, function(x) fmt_pct(x$sens_ed_pct, 1)),
  `Sens BD (%)` = sapply(rr, function(x) fmt_pct(x$spec_bd_pct, 1))
)
kable(rd, digits = 0,
      caption = paste0("Table S3. Comparison of three readings of the lookup table. ",
                       "Sens ED, share of truly effect-dominated repetitions declared usable; ",
                       "Sens BD, share of truly bias-dominated repetitions withheld."))
```

The conservative reading changed `r fmt_pct(cirole$pct_verdict_changed_spec_vs_pool, 2)` of verdicts relative to the pooled reading and `r fmt_pct(cirole$pct_verdict_changed_cons_vs_pool, 2)` relative to the width-specific reading; the maximum absolute within-bin probability gap was `r fmt_num(cirole$max_abs_prob_gap_within_bin, 3)`, and adding CI half-width moved the AUC by only `r fmt_num(cirole$delta_auc_from_ci_width, 5)`.

### S3.3 Nested strategy comparison

Strategies R0 through R4 add one component at a time; C1 and C2 are stand-alone comparators that do not use the calibrated P-value. Every rule is scored twice, once from each viewpoint of Table 4 in the main text, so Table S4 reports eight operating characteristics per rule. The two readings are not mirror images because truth has three levels and the mixed zone is negative in both.

```{r strategy-table}
RULE_LABEL <- c(
  R0 = "R0 uncalibrated P < .05 (no bias correction)",
  R1 = "R1 calibrated P < .05 (layer 1 only)",
  R2 = "R2 layer 1, plus BF point estimate below 0.5, the bias-dominated threshold",
  R3 = "R3 layer 1, plus whole credible interval of BF below 1/3, clear of bias dominance",
  R4 = "R4 layer 1, plus probability of bias dominance below 15%",
  C1 = "C1 BF point estimate below 0.5, the bias-dominated threshold, no layer 1",
  C2 = "C2 whole credible interval of BF below 1/3, clear of bias dominance, no layer 1"
)
# 双视角 10 列。旧版本的 "Spec BD" 列命名有误：它拿到的是
# strat$spec_bd_pct，其实是 BD 视角的 sensitivity（真偏倚主导且被扣下），
# 不是 specificity。R/23 重命名后数字不变、语义对齐。
# Misuse 列删掉：它等于 100 减 NPV(BD)，保留会重复且占宽度。
stopifnot(identical(dual$strategies$rule,
                    sub(" .*", "", sapply(strat, `[[`, "strategy"))))
st <- data.frame(
  # 规则标签：管线里的内部说法换成读者能独立读懂的表述
  # （与正文 Table 4 的 Definition 列同口径，避免两处说法打架）
  Strategy            = unname(RULE_LABEL[dual$strategies$rule]),
  `Declare usable (%)` = fmt_pct(dual$strategies$declare_usable_pct, 1),
  `Sens, ED (%)`      = fmt_pct(dual$strategies$sens_ed, 1),
  `Spec, ED (%)`      = fmt_pct(dual$strategies$spec_ed, 1),
  `PPV, ED (%)`       = fmt_pct(dual$strategies$ppv_ed, 1),
  `NPV, ED (%)`       = fmt_pct(dual$strategies$npv_ed, 1),
  `Sens, BD (%)`      = fmt_pct(dual$strategies$sens_bd, 1),
  `Spec, BD (%)`      = fmt_pct(dual$strategies$spec_bd, 1),
  `PPV, BD (%)`       = fmt_pct(dual$strategies$ppv_bd, 1),
  `NPV, BD (%)`       = fmt_pct(dual$strategies$npv_bd, 1),
  check.names = FALSE
)
kable(st, digits = 0, align = c("l", rep("r", 9)),
      caption = paste0(
        "Table S4. Nested strategy comparison, read from two viewpoints. ",
        "In the effect-dominated view (ED) a positive call is to declare a repetition usable and a positive truth is to be truly effect-dominated; ",
        "in the bias-dominated view (BD) a positive call is to withhold and a positive truth is to be truly bias-dominated. ",
        "Sens, share of truly positive repetitions the rule called positive; Spec, share of truly negative repetitions the rule called negative; ",
        "PPV, share of repetitions the rule called positive that were truly positive; NPV, share the rule called negative that were truly negative. ",
        "The misuse rate reported in Table 4 of the main text is 100 minus NPV in the BD view. Truth has three levels (effect-dominated ",
        kpct(dual$truth_share$effect_dominated, 1), ", mixed ", kpct(dual$truth_share$mixed, 1),
        ", bias-dominated ", kpct(dual$truth_share$bias_dominated, 1),
        "), and the mixed zone is negative in both views, so the views are not mirror images and no pair of columns sums to 100."))
```

### S3.4 Decision flow by stage

```{r flow-table}
fl <- data.frame(
  Stage = sapply(flow, `[[`, "stage"),
  N = sapply(flow, `[[`, "n"),
  `Bias-dominated (n)` = sapply(flow, `[[`, "bd_true_n"),
  `BD share (%)` = sapply(flow, function(x) fmt_pct(x$bd_share_pct, 1))
)
kable(fl, digits = 0,
      caption = "Table S5. Decision flow across layers of the rule.")
```

### S3.5 Verdict versus truth (3 by 3 confusion)

```{r confusion-table}
cf <- data.frame(
  Truth = sapply(confusion3x3, `[[`, "truth"),
  `bias-dominated` = sapply(confusion3x3, `[[`, "bias-dominated"),
  `mixed` = sapply(confusion3x3, `[[`, "mixed"),
  `effect-dominated` = sapply(confusion3x3, `[[`, "effect-dominated"),
  `Row total` = sapply(confusion3x3, `[[`, "row_total")
)
kable(cf, digits = 0,
      caption = "Table S6. Confusion matrix of assigned verdict zone against simulated truth.")
```

Zone accuracy was `r fmt_pct(NUM$part2$zone_accuracy$overall, 1)` overall, `r fmt_pct(NUM$part2$zone_accuracy$bias_dominated, 1)` among bias-dominated, `r fmt_pct(NUM$part2$zone_accuracy$mixed, 1)` among mixed, and `r fmt_pct(NUM$part2$zone_accuracy$effect_dominated, 1)` among effect-dominated repetitions.

### S3.6 Subgroup robustness

```{r subgroup-table}
sg <- data.frame(
  Subgroup = sapply(subg, `[[`, "subgroup"),
  N = sapply(subg, `[[`, "n"),
  `Layer1 pass (%)` = sapply(subg, function(x) fmt_pct(x$layer1_pass_pct, 1)),
  `BD among pass (%)` = sapply(subg, function(x) fmt_pct(x$bd_among_layer1_pct, 1)),
  `Declare usable (%)` = sapply(subg, function(x) fmt_pct(x$declare_usable_pct, 1)),
  `BD among usable (%)` = sapply(subg, function(x) fmt_pct(x$bd_among_declared_pct, 2)),
  `Median half-width` = sapply(subg, `[[`, "median_half_width")
)
kable(sg, digits = c(0, 0, 1, 1, 1, 2, 3),
      caption = paste0("Table S7. Misuse rate across ", length(subg), " pre-specified subgroups."))
```

### S3.7 Discrimination

```{r discrimination-table}
dc <- data.frame(
  Predictor = sapply(disc, `[[`, "predictor"),
  `AUC` = sapply(disc, function(x) fmt_num(x$auc, 3)),
  `LR chi2 note` = sapply(disc, function(x) ifelse(x$note == "", "-", x$note))
)
kable(dc, digits = 0,
      caption = "Table S8. Discrimination (AUC). The area under the curve is read from the bias-dominated view, meaning that withholding a repetition is a positive call and being truly bias-dominated is a positive truth. Figure 8 of the main text reports the same three scores from both viewpoints.")
```

The likelihood-ratio chi-square test for adding credible-interval half-width to BF̂ was χ² = `r sub(".*chi2 = ([0-9.]+).*", "\\1", disc[[3]]$note)`, df = 1, p < .001. The odds ratio per unit increase in CI half-width was `r fmt_num(NUM$part2$or_half_width, 2)`.

### S3.8 Credible-interval width and coverage

The median half-width was `r fmt_num(ciw$median_half_width, 3)` (quartiles `r fmt_num(ciw$quartiles_half[1], 3)`, `r fmt_num(ciw$quartiles_half[3], 3)`). Coverage of the MCMC 95% credible interval was `r fmt_pct(100 * ciw$coverage_overall, 1)` overall, `r fmt_pct(100 * ciw$coverage_narrow, 1)` for narrow intervals, and `r fmt_pct(100 * ciw$coverage_wide, 1)` for wide intervals. Mean half-width by K was `r fmt_num(ciw$mean_half_by_K[1], 3)` for K = 12 and `r fmt_num(ciw$mean_half_by_K[2], 3)` for K = 25. The calibrated P-value at the 0.05 threshold declared `r fmt_pct(100 * l1$pass_rate_overall, 1)` of repetitions passable versus `r fmt_pct(100 * l1$naive_pass_rate, 1)` for the uncalibrated P-value; the pass rate among bias-dominated truth was `r fmt_pct(100 * l1$pass_rate_by_truth$bias_dominated, 1)`, among mixed `r fmt_pct(100 * l1$pass_rate_by_truth$mixed, 1)`, and among effect-dominated `r fmt_pct(100 * l1$pass_rate_by_truth$effect_dominated, 1)`.

---

## S4. Part III supplementary: case-study estimation

### S4.1 Negative-control empirical null

The jointly estimated empirical null had center μ̂_B = `r fmt_num(cases[[1]]$mu_b, 3)` and scale σ̂_B = `r fmt_num(cases[[1]]$sigma_b, 3)`; the Shapiro-Wilk test of the `r fmt_num(nc$K, 0)` negative-control residuals gave p = `r fmt_num(nc$shapiro_p, 3)`, and the calibrated log relative risks of the controls ranged from `r fmt_num(nc$logrr_min, 3)` to `r fmt_num(nc$logrr_max, 3)`.

### S4.2 Full case estimation

```{r case-table}
cs <- data.frame(
  Question = c(cases[[1]]$exposure, cases[[2]]$exposure),
  `Uncal RR` = c(cases[[1]]$rr_uncal, cases[[2]]$rr_uncal),
  `Cal RR` = c(cases[[1]]$rr_cal, cases[[2]]$rr_cal),
  `Cal P` = c(cases[[1]]$cal_p, cases[[2]]$cal_p),
  `BFhat` = c(cases[[1]]$bf, cases[[2]]$bf),
  `BF lo` = c(cases[[1]]$bf_lo, cases[[2]]$bf_lo),
  `BF hi` = c(cases[[1]]$bf_hi, cases[[2]]$bf_hi),
  `p bias-dom` = c(cases[[1]]$p_bias_dominated, cases[[2]]$p_bias_dominated),
  Verdict = c(cases[[1]]$verdict, cases[[2]]$verdict)
)
kable(cs, digits = c(0, 3, 3, 4, 3, 3, 3, 3, 0),
      caption = "Table S9. Full estimation output for the two clinical questions.")
```

### S4.3 Sensitivity analyses (question 1)

Leave-one-out across the `r fmt_num(nc$K, 0)` negative controls gave BF̂ ranging from `r fmt_num(sens$loo_bf_min, 3)` to `r fmt_num(sens$loo_bf_max, 3)` (fall-back BF̂ = `r fmt_num(sens$loo_fall_bf, 3)`). Replacement of the Super Learner library gave BF̂ = `r fmt_num(sens$sl_bf[1], 3)` (`r fmt_num(sens$sl_bf_lo[1], 3)` to `r fmt_num(sens$sl_bf_hi[1], 3)`), `r fmt_num(sens$sl_bf[2], 3)` (`r fmt_num(sens$sl_bf_lo[2], 3)` to `r fmt_num(sens$sl_bf_hi[2], 3)`), and `r fmt_num(sens$sl_bf[3], 3)` (`r fmt_num(sens$sl_bf_lo[3], 3)` to `r fmt_num(sens$sl_bf_hi[3], 3)`) under the three alternative libraries. The Gaussian-process propensity-score perturbation gave calibrated relative risk `r fmt_num(sens$gp_cal_rr[1], 3)` to `r fmt_num(sens$gp_cal_rr[2], 3)` and BF̂ `r fmt_num(sens$gp_bf[1], 3)` to `r fmt_num(sens$gp_bf[2], 3)`. All sensitivity analyses preserved the "not usable as effect evidence" verdict for question 1 and the "competitive, no verdict" zone for question 2.

---

## S5. Software and reproducibility

The estimator is implemented in the R package EmpiricalCalibration [@schuemie2018empirical]; the BF extension is provided by the accompanying R code archived with this manuscript. The analysis pipeline reads MIMIC IV and MIMIC IV ECG under approved data-use agreements; the derived negative-control and estimation tables, the simulation frame, and the single JSON of reported numbers are released with the manuscript repository. The companion interactive explainer consumes the same JSON so that every figure and table in the manuscript and the explainer are guaranteed to agree.
